---
title: "20230327_MLMProject_ZC"
output: html_document
date: "2023-03-27"
---
# SET UP
```{r}
# load packages
library(tidyverse) #data wrangling
library(dplyr) # data wrangling
library(ggpubr) # plotting
library(readr) # read_csv
library(psych) # psych statistics - MSSD
library(lme4) # mixed models
library(effectsize) #effect size functions
library(lubridate) #time variables
```

# EMA
```{r}
filename <- "20230321_EMAData.csv"

EMAdata <- read_csv(filename)
```
```{r}
filename <- "20230327_betweenday.csv"

betweenday <- read_csv(filename)
```

```{r}
#transform variables to factor
EMAdata$id <- as.factor(EMAdata$id)
EMAdata$event <- as.factor(EMAdata$event) 
levels(EMAdata$event) <- c("Baseline", "Session 4", "Post")
EMAdata$enrollmentgroup <- as.factor(EMAdata$enrollmentgroup) 
levels(EMAdata$enrollmentgroup) <- c("Waitlist Control", "Immediate Enrollment", "Group")
```

## Descriptive Statistics
```{r}
#remove group 2
EMAdata <- filter(EMAdata, enrollmentgroup != "Group")
```

### Missing Data

Missing Instances
```{r}
#create dataframe of missing instances (completed = 0)
missingdata <- filter(EMAdata, completed == '0')
```

```{r}
#create table of n missing: event x group
missing_instance <- matrix(0, 4,4)

colnames(missing_instance) <- c("Baseline", "Session4", "Post", "Total")
rownames(missing_instance) <- c("WLC", "IE", "Group", "Total")
```

```{r}
missing_instance[1,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "0"))
missing_instance[2,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "1"))
missing_instance[3,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "2"))
missing_instance[1,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "0"))
missing_instance[2,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "1"))
missing_instance[3,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "2"))
missing_instance[1,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "0"))
missing_instance[2,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "1"))
missing_instance[3,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "2"))

missing_instance[1,4] <- length(which(missingdata$enrollmentgroup == "0"))
missing_instance[2,4] <- length(which(missingdata$enrollmentgroup == "1"))
missing_instance[3,4] <- length(which(missingdata$enrollmentgroup == "2"))

missing_instance[4,1] <- length(which(missingdata$event == "baseline"))
missing_instance[4,2] <- length(which(missingdata$event == "session4"))
missing_instance[4,3] <- length(which(missingdata$event == "post"))
missing_instance[4,4] <- missing_instance[4,1] + missing_instance[4,2] + missing_instance[4,3]
```

Missing Items
```{r}
#create dataframe of missing items (completed =1, N/A on at least 1 item)
missingitems <- filter(EMAdata, completed == "1" & (is.na(positive_score) | is.na(negative_score))) #n=9
```

```{r}
#impute missing items using the average of other neg/pos items
## create function

#function inputs: ID, EVENT, INSTANCE
itemavg <- function(ID, EVENT, INSTANCE){

  # row from id, event, instance
  r <- which(EMAdata$id == ID & EMAdata$event == EVENT & EMAdata$instance == INSTANCE)  
  
  positems <- c()
  negitems <- c()

  #positive items: 1, 3, 5, 9, 10, 12, 14, 16, 17, 19
  positems <- append(positems, as.integer(EMAdata[r,15])) #posq1
  positems <- append(positems, as.integer(EMAdata[r,17])) #posq3
  positems <- append(positems, as.integer(EMAdata[r,19]))#posq5
  positems <- append(positems, as.integer(EMAdata[r,23]))#posq9
  positems <- append(positems, as.integer(EMAdata[r,24])) #posq10
  positems <- append(positems, as.integer(EMAdata[r,26])) #posq12
  positems <- append(positems, as.integer(EMAdata[r,28])) #posq14
  positems <- append(positems, as.integer(EMAdata[r,30])) #posq16
  positems <- append(positems, as.integer(EMAdata[r,31])) #posq17
  positems <- append(positems, as.integer(EMAdata[r,33])) #posq19

  posavg <- mean(positems, na.rm=TRUE)

  #negative items: 2, 4, 6, 7, 8, 11, 13, 15, 18, 20
  negitems <- append(negitems,  as.integer(EMAdata[r,16])) #negq2
  negitems <- append(negitems, as.integer(EMAdata[r,18])) #negq4
  negitems <- append(negitems, as.integer(EMAdata[r,20])) #negq6
  negitems <- append(negitems, as.integer(EMAdata[r,21])) #negq7
  negitems <- append(negitems, as.integer(EMAdata[r,22])) #negq8
  negitems <- append(negitems, as.integer(EMAdata[r,25])) #negq11
  negitems <- append(negitems, as.integer(EMAdata[r,27])) #negq13
  negitems <- append(negitems, as.integer(EMAdata[r,29])) #negq15
  negitems <- append(negitems, as.integer(EMAdata[r,32])) #negq18
  negitems <- append(negitems, as.integer(EMAdata[r,34])) #negq20
  
  negavg <- mean(negitems, na.rm=TRUE)
  
  #output posavg, negavg
  out <- list(r, posavg, negavg)
  
  return(out)
}  
```

```{r}
#add all positive items
posscore <- function(r){
  score <- as.integer(EMAdata[r, 15]) + as.integer(EMAdata[r, 17]) + 
    as.integer(EMAdata[r, 19]) + as.integer(EMAdata[r, 23]) +
    as.integer(EMAdata[r, 24]) + as.integer(EMAdata[r, 26]) + 
    as.integer(EMAdata[r, 28]) + as.integer(EMAdata[r, 30]) + 
    as.integer(EMAdata[r, 31]) + as.integer(EMAdata[r, 33])
  
  return(score)
}

#add all negative items  
negscore <- function(r){
  score <- as.integer(EMAdata[r, 16]) + as.integer(EMAdata[r, 18]) + 
    as.integer(EMAdata[r, 20]) + as.integer(EMAdata[r, 21]) +
    as.integer(EMAdata[r, 22]) + as.integer(EMAdata[r, 25]) + 
    as.integer(EMAdata[r, 27]) + as.integer(EMAdata[r, 29]) + 
    as.integer(EMAdata[r, 32]) + as.integer(EMAdata[r, 34])
  
  return(score)
}
  
```

```{r}
#12015 session 4:8
out<- itemavg("12015", "Session 4", 8)
EMAdata[23, 17] <- out[2] #posq3 
EMAdata[23, 23] <- out[2] #posq9 
EMAdata[23,11] <- posscore(23) #positive score

#12015 session4: 12
out<- itemavg("12015", "Session 4", 12)
EMAdata[27, 25] <- out[3] #negq11
EMAdata[27,12] <- negscore(27) #negative score

#12017 post: 3
out <- itemavg("12017", "Post", 3)
EMAdata[123, 30] <- out[2] #posq16
EMAdata[123, 11] <- posscore(123)

#12027 baseline:2
out <- itemavg("12027", "Baseline", 2)
EMAdata[542, 28] <- out[2] #posq14
EMAdata[542, 11] <- posscore(542)

#12027 session4:12
out <- itemavg("12027", "Session 4", 12)
EMAdata[567, 22] <- out[3] #neg8
EMAdata[567, 27] <- out[3] #neg13
EMAdata[567,12] <- negscore(567)

#12033 baseline:15
out<- itemavg("12033", "Baseline", 15)
EMAdata[810, 24] <- out[2] #pos10
EMAdata[810, 11] <- posscore(810)

#12033 session4:11
out<- itemavg("12033", "Session 4", 11)
EMAdata[821, 30] <- out[2] #posq16
EMAdata[821, 11] <- posscore(821)

#12039 post:2
out<- itemavg("12039", "Post", 2)
EMAdata[1097, 31] <- out[2] #posq17
EMAdata[1097,11] <- posscore(1097)

#12039 post:3
out<- itemavg("12039", "Post", 3)
EMAdata[1098, 28] <- out[2] #posq14
EMAdata[1098, 34] <- out[3] #neg20
EMAdata[1098, 11] <- posscore(1098)
EMAdata[1098, 12] <- negscore(1098)
```

```{r}
#check missing items (completed =1, N/A on at least 1 item)
missingitems <- filter(EMAdata, completed == "1" & (is.na(positive_score) | is.na(negative_score))) #n=9

view(missingitems)
```

### Survey Completion

Compliance Rate

```{r}
#create dataframe of id x completion
compliance_rate <- matrix(0, 35, 3)

colnames(compliance_rate) <- c("Group", "Completed", "ComplianceRate")
id_names <- c("12015", "12016", "12017", "12018", "12019", "12020","12021", "12022", "12023", "12024","12025", "12026", "12027", "12028", "12029", "12030", "12031", "12032", "12033", "12034", "12035", "12036", "12037", "12038", "12039", "12040", "12041","12043", "12044", "12045", "12048","12050", "12055","12056", "12057")
rownames(compliance_rate) <- id_names
```

```{r}
#fill matrix col 1 with enrollment group
r = 1
n = 1

for(n in 1:35){
  ind <- which(EMAdata$id == id_names[n])
  compliance_rate[r,1] <- EMAdata$enrollmentgroup[ind[1]]
  r <- r + 1
  n <- n + 1
}

```

```{r}
#fill matrix col 2 with n completed by id
r = 1
n = 1

for(n in 1:35){
  compliance_rate[r,2] <- length(which(EMAdata$id == id_names[n]))
  r <- r + 1
  n <- n + 1
  }
```

```{r}
#fill matrix col 3 with compliance rate
r = 1
n = 1

for (n in 1:35){
  compliance_rate[r, 3] <- compliance_rate[r, 2]/45
  r <- r + 1
  n <- n + 1
}
```

```{r}
#output min and max values
max <- which(compliance_rate == max(compliance_rate[,3]), arr.ind=TRUE)
min <- which(compliance_rate == min(compliance_rate[,3]), arr.ind=TRUE)

print(paste("Maximum compliance rate:", compliance_rate[max[1]]))
print(paste("Minimum compliance rate:", compliance_rate[min]))
```

### Mean (SD) Table
```{r}
summarise_at(EMAdata, vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#group by event
EMAdata%>%
  group_by(event) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#group by enrollment
EMAdata%>%
  group_by(enrollmentgroup) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```
### Normality & Distribution
```{r}
#create grouped boxplot - positive

ggplot(EMAdata, aes(x=event, y=positive_score, fill=enrollmentgroup)) + scale_fill_manual(values = c("lightblue3", "dodgerblue4")) + geom_boxplot() + xlab("event") + ylab("score") + ggtitle("PANAS Positive Score x Group")
```

```{r}
#create grouped boxplot - negative

ggplot(EMAdata, aes(x=event, y=negative_score, fill=enrollmentgroup)) + 
  scale_fill_manual(values = c("lightblue3", "dodgerblue4")) +
  geom_boxplot() + xlab("event") + ylab("score") +
  ggtitle("PANAS Negative Score x Group")
```
## Affect Dynamic Variables

```{r}
withinday <- EMAdata[,c('id', 'event', 'day', 'instance', 'enrollmentgroup', 'time_interval', 'positive_score', 'negative_score', "positive_SD","negative_SD")]
```

```{r}
#convert time_interval from h:m:s to minutes
withinday$time_interval_min <- hour(withinday$time_interval)*60 + minute(withinday$time_interval)

withinday <- withinday %>%
  relocate("time_interval_min", .after="time_interval")
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Negative")
```

```{r}
#ASD = positive_SD[i] / time_interval[i]
pos_ASD <- vector()

for(i in 1:1427){
  pos_ASD <- append(pos_ASD, c(withinday$positive_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(pos_ASD)
```

```{r}
#create positive ASD variable
withinday$positive_ASD <- pos_ASD
```

```{r}
neg_ASD <- vector()

for(i in 1:1427){
  neg_ASD <- append(neg_ASD, c(withinday$negative_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(neg_ASD)
```

```{r}
#create negative ASD variable
withinday$negative_ASD <- neg_ASD
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Negative")
```

```{r}
betweenday <- filter(betweenday, betweenday$id != "12046")
  
#split betweenday df by event
betweenday_baseline <- filter(betweenday, betweenday$event == "1")
betweenday_session4 <- filter(betweenday, betweenday$event == "2")
betweenday_post <- filter(betweenday, betweenday$event == "3")
```

```{r}
#remove group 2 & 12046 (no between data)
id_names <- c("12015", "12016", "12017", "12018", "12019", "12020","12021", "12022", "12023", "12024","12025", "12026", "12027", "12028", "12029", "12030", "12031", "12032", "12033", "12034", "12035", "12036", "12037", "12038", "12039", "12040", "12041","12043", "12044", "12045", "12050", "12057")

length(id_names)

id_group <- c("1", "1", "0", "1", "0", "1", "0", "0", "1", "1", "0", "1", "0", "0", "0", "0", "1", "1", "1", "0", "1", "0", "0", "1", "1", "0", "0", "1", "0", "1","1", "1")

length(id_group)
```

### Mean 

```{r}
# initialize MEAN matrix
ema_mean <- matrix(0, 32, 7)

colnames(ema_mean) <- c("group", "posbaseline", "possession4", "pospost", "negbaseline", "negsession4", "negpost")
rownames(ema_mean) <- id_names

for(i in 1:32){
  ema_mean[i,1] <- id_group[i]
}
```

```{r}
#fill positive mean
## baseline

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_baseline$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_baseline$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,2] <- mean(vec)
}

## session4

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_session4$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_session4$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,3] <- mean(vec)
}

## post

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_post$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_post$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,4] <- mean(vec)
}
```

```{r}
#fill negative mean
## baseline

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_baseline$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_baseline$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,5] <- mean(vec)
}

## session4

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_session4$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_session4$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,6] <- mean(vec)
}

## post

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_post$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_post$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_mean[i,7] <- mean(vec)
}
```

```{r}
ema_mean <- as.data.frame(ema_mean)
ema_mean$posbaseline <- as.numeric(ema_mean$posbaseline)
ema_mean$possession4 <- as.numeric(ema_mean$possession4)
ema_mean$pospost <- as.numeric(ema_mean$pospost)
ema_mean$negbaseline <- as.numeric(ema_mean$negbaseline)
ema_mean$negsession4 <- as.numeric(ema_mean$negsession4)
ema_mean$negpost <- as.numeric(ema_mean$negpost)
```

```{r}
#group by group
ema_mean%>%
  group_by(group) %>%
  summarise_at(vars(posbaseline, possession4, pospost, negbaseline, negsession4, negpost, ), list(mean=mean, sd=sd), na.rm=TRUE)
```
### Variability (SD)

```{r}
# initialize SD matrix
ema_sd <- matrix(0, 32, 7)

colnames(ema_sd) <- c("group", "posbaseline", "possession4", "pospost", "negbaseline", "negsession4", "negpost")
rownames(ema_sd) <- id_names

for(i in 1:32){
  ema_sd[i,1] <- id_group[i]
}
```

```{r}
#fill positive mean
## baseline

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_baseline$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_baseline$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,2] <- sd(vec)
}

## session4

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_session4$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_session4$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,3] <- sd(vec)
}

## post

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_post$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_post$positive_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,4] <- sd(vec)
}
```

```{r}
#fill negative mean
## baseline

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_baseline$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_baseline$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,5] <- sd(vec)
}

## session4

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_session4$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_session4$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,6] <- sd(vec)
}

## post

for(i in 1:32){
  name <- id_names[i]
  index <- which(betweenday_post$id == name)
  
  vec <- c()
  len <- length(index)

  for(x in 1:len){
    n <- betweenday_post$negative_mean[index[x]]
    vec <- c(vec, n)
  }
  
  ema_sd[i,7] <- sd(vec)
}
```

```{r}
ema_sd <- as.data.frame(ema_sd)
ema_sd$posbaseline <- as.numeric(ema_sd$posbaseline)
ema_sd$possession4 <- as.numeric(ema_sd$possession4)
ema_sd$pospost <- as.numeric(ema_sd$pospost)
ema_sd$negbaseline <- as.numeric(ema_sd$negbaseline)
ema_sd$negsession4 <- as.numeric(ema_sd$negsession4)
ema_sd$negpost <- as.numeric(ema_sd$negpost)
```

```{r}
#group by group
ema_sd%>%
  group_by(group) %>%
  summarise_at(vars(posbaseline, possession4, pospost, negbaseline, negsession4, negpost, ), list(mean=mean, sd=sd), na.rm=TRUE)
```

### Intertia (Auto-Correlation)

### Affect Instability (MSSD)

```{r}
# initialize MSSD matrix
ema_mssd <- matrix(0, 32, 7)

colnames(ema_mssd) <- c("group", "posbaseline", "possession4", "pospost", "negbaseline", "negsession4", "negpost")
rownames(ema_mssd) <- id_names

for(i in 1:32){
  ema_mssd[i,1] <- id_group[i]
}
```

```{r}
#fill positive mssd
## baseline
mssd <- mssd(betweenday_baseline$positive_mean, group = betweenday_baseline$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
  ema_mssd[i,2] <- mssd[i]
}

## session4
mssd <- mssd(betweenday_session4$positive_mean, group = betweenday_session4$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
  ema_mssd[i,3] <- mssd[i]
}

## post
mssd <- mssd(betweenday_post$positive_mean, group = betweenday_post$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
  ema_mssd[i,4] <- mssd[i]
}
```

```{r}
#fill negative mssd
## baseline
mssd <- mssd(betweenday_baseline$negative_mean, group = betweenday_baseline$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
  ema_mssd[i,5] <- mssd[i]
}

## session 4
mssd <- mssd(betweenday_session4$negative_mean, group = betweenday_session4$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
 ema_mssd[i,6] <- mssd[i]
}

## post
mssd <- mssd(betweenday_post$negative_mean, group = betweenday_post$id, lag = 1, na.rm = TRUE)

for(i in 1:32){
  ema_mssd[i,7] <- mssd[i]
}
```

```{r}
ema_mssd <- as.data.frame(ema_mssd)
ema_mssd$posbaseline <- as.numeric(ema_mssd$posbaseline)
ema_mssd$possession4 <- as.numeric(ema_mssd$possession4)
ema_mssd$pospost <- as.numeric(ema_mssd$pospost)
ema_mssd$negbaseline <- as.numeric(ema_mssd$negbaseline)
ema_mssd$negsession4 <- as.numeric(ema_mssd$negsession4)
ema_mssd$negpost <- as.numeric(ema_mssd$negpost)
```

```{r}
#group by group
ema_mssd%>%
  group_by(group) %>%
  summarise_at(vars(posbaseline, possession4, pospost, negbaseline, negsession4, negpost, ), list(mean=mean, sd=sd), na.rm=TRUE)
```

### Acute Increase (PAC)

## Preliminary Analyses
mixed model w REML
fixed effects: enrollment group, event, event*enrollmentgroup
random effects: id

calculated in excel - kept running into issues doing it in R
```{r}
filename <- "20230325_betweenday.csv"

betweenday <- read_csv(filename)
```

### x Events
```{r}
lmm_emapos <- lmer(positive_mean ~ enrollmentgroup * event + (1|id), data=betweenday)

summary(lmm_emapos)
anova(lmm_emapos) 

eta_squared(lmm_emapos, partial=FALSE)
```

```{r}
lmm_emaneg <- lmer(negative_mean ~ enrollmentgroup * event + (1|id), data=betweenday)

summary(lmm_emaneg)
anova(lmm_emaneg) 

eta_squared(lmm_emaneg, partial=FALSE)
```

#### Plots
```{r}
# Positive Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panaspos <- confint(lmm_emapos, method="boot", nsim=10)
ci_panaspos <- tibble::rownames_to_column(data.frame(ci_panaspos), "Term") 
colnames(ci_panaspos)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panaspos)

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=positive_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Timepoints") + 
  labs(y = "positive affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lmm_emapos)[1], #Regression Line (RL).
              slope=fixef(lmm_emapos)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panaspos$`CI 97.5%`[3], 
              slope=ci_panaspos$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panaspos$`CI 2.5%`[3], 
              slope=ci_panaspos$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL
  
panaspos_ggplot
```

```{r}
# Negative Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panasneg <- confint(lmm_emaneg, method="boot", nsim=10)
ci_panasneg <- tibble::rownames_to_column(data.frame(ci_panasneg), "Term") 
colnames(ci_panasneg)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panasneg)

#Plot model 
panasneg_ggplot<- ggplot(data= betweenday, 
       aes(y=negative_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Timepoints") + 
  labs(y = "negative affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lmm_emaneg)[1], #Regression Line (RL).
              slope=fixef(lmm_emaneg)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panasneg$`CI 97.5%`[3], 
              slope=ci_panasneg$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panasneg$`CI 2.5%`[3], 
              slope=ci_panasneg$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL

panasneg_ggplot
```

### x Days
```{r}
betweenday$day <- as.factor(betweenday$day)
levels(betweenday$day) <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")
```

```{r}
lmm_emapos <- lmer(positive_mean ~ enrollmentgroup * day + (1|id), data=betweenday)

anova(lmm_emapos) 
summary(lmm_emapos)

eta_squared(lmm_emapos, partial = FALSE)
```

```{r}
lmm_emaneg <- lmer(negative_mean ~ enrollmentgroup * day + (1|id), data=betweenday)

anova(lmm_emaneg) 
summary(lmm_emaneg)

eta_squared(lmm_emaneg, partial = FALSE)
```

#### Plots
```{r}
# Positive Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=positive_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Days") + 
  labs(y = "positive affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) 
 
panaspos_ggplot
```

```{r}
# Negative Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=negative_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Days") + 
  labs(y = "negative affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels)

panaspos_ggplot
```

### x Instances
```{r}
EMAdata$instance <- as.factor(EMAdata$instance)
levels(EMAdata$instance) <- c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15")
```

```{r}
lmm_emapos <- lmer(positive_score ~ enrollmentgroup * instance + (1|id), data= EMAdata)

anova(lmm_emapos) 
summary(lmm_emapos)
```

```{r}
lme_emaneg <- lmer(negative_score ~ enrollmentgroup * instance + (1|id), data= EMAdata)

anova(lme_emaneg) 
summary(lme_emaneg)
```

#### Plots
```{r}
# Positive Affect plot @ Baseline
EMA_baseline <- filter(EMAdata, event == "Baseline")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_baseline, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Baseline") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Positive Affect plot @ Session 4
EMA_session4 <- filter(EMAdata, event == "Session 4")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_session4, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Session4") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Positive Affect plot @ Post
EMA_post <- filter(EMAdata, event == "Post")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_post, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Post") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Negative Affect plot @ Baseline

instance_neg<- ggplot(data= EMA_baseline, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Baseline") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```

```{r}
# Negative Affect plot @ Session 4

instance_neg<- ggplot(data= EMA_session4, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Session 4") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```

```{r}
# Negative Affect plot @ Post

instance_neg<- ggplot(data= EMA_post, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Post") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```


