---
title: "MBMS_DataAnalysis_Phase1"
output: html_document
---
Open necessary packages
```{r}
install.packages("lme4")
library(tibble)
library(dplyr)
library(data.table)
library(lme4)
library(nlme)
```

Filter REDCap data for desired IDs. In this round, 12002 is excluded because the child was TD, and 12009 did not complete the study. 
```{r}
PrePostData_Filter <- PrePostData %>%
  filter(Record_ID != "12002" & Record_ID != "12009")
```
I will remove any missing data from the data set. 
```{r}
PrePostData_Filter<- PrePostData_Filter %>%
  filter(!is.na(PANAS_Start_Pos),
         !is.na(PANAS_Start_Neg),
         !is.na(IOS_Pre),
         !is.na(PANAS_End_Pos),
         !is.na(PANAS_End_Neg),
         !is.na(IOS_Post))
```

For the first analysis, I will only be using Pre data points. So, I will filter for only this data.Named _An1 for analysis 1.
```{r}
PreOnlyData_An1 <- PrePostData_Filter %>%
  select(Record_ID, Session, PANAS_Start_Pos, PANAS_Start_Neg, IOS_Pre)

as_tibble(PreOnlyData_An1)
```

Analysis 1 will begin with descriptive statistics of the three variables: PANAS_Start_Pos, PANAS_Start_Neg, IOS_Pre
```{r}
PreOnlyData_An1 %>%
  sapply (mean,na.rm=TRUE)
PreOnlyData_An1 %>%
  sapply (sd,na.rm=TRUE) 
PreOnlyData_An1 %>%
  sapply (range, na.rm=TRUE)

```

Beginning with PANAS_Start_Pos. The data is nested by timepoint 1,2,3,4,5,6,7. First, plot PANAS_Start_Pos by timepoint(Session). 
```{r}
PANAS_Pos_lm=lm(PANAS_Start_Pos ~ Session, data=PreOnlyData_An1)
summary(PANAS_Pos_lm)
```
## Output shows the following
## Residual standard error =8.898, df(80)
## R2= 0.0024, Adjusted R2= -0.010
## F(1,80)=0.1954, p=0.6597, P>.05

Add analysis of group effects by individuals (Record_ID)
```{r}
PANAS_Pos_mixed =lmer(PANAS_Start_Pos ~ Session + (1|Record_ID), data=PreOnlyData_An1)

summary(PANAS_Pos_mixed)

confint(PANAS_Pos_mixed)
```

Now run the models with PANAS_Start_Neg.
```{r}
PANAS_Neg_lm=lm(PANAS_Start_Neg ~ Session, data=PreOnlyData_An1)
summary(PANAS_Neg_lm)
```
## Output shows the following
## Residual standard error =4.471, df(80)
## R2= 0.118, Adjusted R2= 0.1069
## F(1,80)=10.7, p=0.001583, P<.05

Add analysis of group effects by individuals (Record_ID)
```{r}
PANAS_Neg_mixed=lmer(PANAS_Start_Neg ~ Session + (1|Record_ID), data=PreOnlyData_An1)
summary(PANAS_Neg_mixed)
confint(PANAS_Neg_mixed)
```

Now run the models with IOS_Pre.
```{r}
IOS_lm=lm(IOS_Pre ~ Session, data=PreOnlyData_An1)
summary(IOS_lm)
```
## Output shows the following
## Residual standard error =1.726, df(80)
## R2= 0.2795, Adjusted R2= 0.2705
## F(1,80)=31.03, p=0.00, P<.05
Add analysis of group effects by individuals (Record_ID)
```{r}
IOS_mixed=lmer(IOS_Pre ~ Session + (1|Record_ID), data=PreOnlyData_An1)
summary(IOS_mixed)
confint(IOS_mixed)
```

Read in Baseline, Post, Follow-Up Data
Variables are:
    Ryff's Wellbeing Scale -> Ryff_Total
    Sleep Inventory -> Sleep_Total
    Social Connectedness Scale -> SCS_Total
    Percieved Stress Scale -> PSS_Total
    Mindfulness Inventory -> MI_Total
    Flow State Scale -> Flow_Total
    Meaningfulness of Songwriting -> MofS_Total
    BDI Raw Score Rounded -> BDI_RawScore
    BAI Raw Score Rounded -> BAI_RawScore
  
```{r}
library(readr)
MBMS_BaselinePostFollowup <- read_csv("MBMS-BaselinePostFollowup.csv")
View(MBMS_BaselinePostFollowup)
```

Filter data by event: Baseline, Post, Follow-up
```{r}
MBMS_Baseline <- MBMS_BaselinePostFollowup %>%
  filter(Event=="Baseline")%>%
  select(Ryff_Total, Sleep_Total, SCS_Total, PSS_Total, MI_Total,BDI_RawScore, BAI_RawScore)

MBMS_Post <- MBMS_BaselinePostFollowup %>%
  filter(Event=="Post")%>%
  select(Ryff_Total, Sleep_Total, SCS_Total, PSS_Total, MI_Total,BDI_RawScore, BAI_RawScore)

MBMS_Followup <-MBMS_BaselinePostFollowup %>%
  filter(Event=="Follow-Up")%>%
  select(Ryff_Total, Sleep_Total, SCS_Total, PSS_Total, MI_Total,BDI_RawScore, BAI_RawScore)
```


Run descriptive statistics at each timepoint.
1. Baseline
```{r}
MBMS_Baseline %>%
  sapply (mean,na.rm=TRUE)
MBMS_Baseline %>%
  sapply (sd,na.rm=TRUE) 
MBMS_Baseline %>%
  sapply (range, na.rm=TRUE)
```
2. Post
```{r}
MBMS_Post %>%
  sapply (mean,na.rm=TRUE)
MBMS_Post %>%
  sapply (sd,na.rm=TRUE) 
MBMS_Post %>%
  sapply (range, na.rm=TRUE)
```
3. Follow-Up
```{r}
MBMS_Followup %>%
  sapply (mean,na.rm=TRUE)
MBMS_Followup %>%
  sapply (sd,na.rm=TRUE) 
MBMS_Followup %>%
  sapply (range, na.rm=TRUE)
```

##Mixed Model
##  Level 1: y = B0 + B1(Session) + e
##  Level 2: B0 = g0 + u0 
##  Mixed Model: y = g0 + B1(Session) + u0 + e
  
##  Where g0 -> average score across time points
##        u0 -> random deviation from the intercept for each time point
##        B1 -> average regression coefficient for change associated with Session 
##        e -> individual level one error

Convert Event to Timepoint
  Baseline -> 1
  Post -> 2
  Follow-up -> 3
 
```{r}
Baseline <- "Baseline"
Post <- "Post"
Follow-Up <- "Follow-Up"

attach(MBMS_BaselinePostFollowup)
Time [Event==Baseline]<-1
Time [Event==Post]<-2
Time [Event==Follow-Up]<-3

MBMS_BaselinePostFollowup <- MBMS_BaselinePostFollowup%>%
  mutate(Time=

```

  


  
  
Run mixed models.   
```{r}
MBMS_BaselinePostFollowup %>%
  model_ryff = lme(fixed = Ryff_Total~)
```


```{r}
