---
title: "20230325_NEAREPORT_ZC"
output: html_document
date: '2023-03-25'
---
# SET UP 
```{r}
# load packages
library(tidyverse) #data wrangling
library(dplyr) # data wrangling
library(ggpubr) # plotting
library(readr) # read_csv
library(psych) # psych statistics - MSSD
library(lme4) # mixed models
library(rstatix) #outlier() function
library(MASS) #boxcox function
library(EMAtools) #effect size functions
library(lubridate) #time variables
```

# OUTCOME MEAURES
```{r}
# load datasets
filename <- "20230325_phase2data.csv"
outcomedata <- read_csv(filename)
```
Coded Variables
** Event **
1 - Baseline
2 - Post
3 - Followup

** Group **
0 - Control
1 - Treatment

```{r}
outcomedata <- filter(outcomedata, group != "2") # remove Group

outcomedata$id <- factor(outcomedata$id) 
outcomedata$event <- factor(outcomedata$event, labels = c("baseline", "post", "followup"))
outcomedata$group <- factor(outcomedata$group, labels = c("control", "treatment"))
```

## Descriptive Statistics
### Mean (SD) Table
```{r}
describe(outcomedata[,4:9])
```
```{r}
#baseline statistics
outcomedata_baseline <- filter(outcomedata, event == "baseline")

outcomedata_baseline %>%
  group_by(group)%>%
  get_summary_stats(type = "mean_sd")

outcomedata_baseline%>%
  get_summary_stats(type = "mean_sd")
```
```{r}
#BDI clinical significance (score >= 20)
bdi_wlc_sig <- which(outcomedata_baseline$bdi >=20 & outcomedata_baseline$group == 0) #wlc group
bdi_wlc_n <- length(bdi_wlc_sig)
bdi_wlc_perc <- (bdi_wlc_n/16) * 100
bdi_wlc_perc

bdi_ie_sig <- which(outcomedata_baseline$bdi >=20 & outcomedata_baseline$group == 1) #ie group
bdi_ie_n <- length(bdi_ie_sig)
bdi_ie_perc <- (bdi_ie_n/17) * 100
bdi_ie_perc

((bdi_wlc_n + bdi_ie_n)/33)*100 #total

#BAI clinical significance (score >= 16)
bai_wlc_sig <- which(outcomedata_baseline$bai >=16 & outcomedata_baseline$group == 0) #wlc group
bai_wlc_n <- length(bai_wlc_sig)
bai_wlc_perc <- (bai_wlc_n/16) * 100
bai_wlc_perc

bai_ie_sig <- which(outcomedata_baseline$bai >=16 & outcomedata_baseline$group == 1) #ie group
bai_ie_n <- length(bai_ie_sig)
bai_ie_perc <- (bai_ie_n/17) * 100
bai_ie_perc

((bai_wlc_n + bai_ie_n)/33)*100 #total
```

```{r}
#checkfor significant differences
t.test(ryff ~group, var.equal=TRUE, data=outcomedata_baseline)
t.test(pss ~group, var.equal=TRUE, data=outcomedata_baseline)
t.test(bdi ~group, var.equal=TRUE, data=outcomedata_baseline)
t.test(bai ~group, var.equal=TRUE, data=outcomedata_baseline)
t.test(fmqi ~group, var.equal=TRUE, data=outcomedata_baseline)
t.test(cerq ~group, var.equal=TRUE, data=outcomedata_baseline)
```

### Normality & Distributions
```{r}
#create boxplots
ggboxplot(outcomedata, x = "event", y = "ryff", add = "point")


ggboxplot(outcomedata, x = "event", y = "pss", add = "point")


ggboxplot(outcomedata, x = "event", y = "bdi", add = "point")


ggboxplot(outcomedata, x = "event", y = "bai", add = "point")


ggboxplot(outcomedata, x = "event", y = "fmqi", add = "point")


ggboxplot(outcomedata, x = "event", y = "cerq", add = "point")

```
```{r}
#create qq plots
ggqqplot(outcomedata, "ryff", facet.by = "event")

ggqqplot(outcomedata, "pss", facet.by = "event")

ggqqplot(outcomedata, "bdi", facet.by = "event")

ggqqplot(outcomedata, "bai", facet.by = "event")

ggqqplot(outcomedata, "fmqi", facet.by = "event")

ggqqplot(outcomedata, "cerq", facet.by = "event")
```
```{r}
#check for outliers
outcomedata %>%
  group_by(event) %>%
  identify_outliers(ryff)

outcomedata %>%
  group_by(event) %>%
  identify_outliers(pss)

outcomedata %>%
  group_by(event) %>%
  identify_outliers(bdi)

outcomedata %>%
  group_by(event) %>%
  identify_outliers(bai)

outcomedata %>%
  group_by(event) %>%
  identify_outliers(fmqi)

outcomedata %>%
  group_by(event) %>%
  identify_outliers(cerq)
```
no extreme outliers

```{r}
#check normality assumption
outcomedata %>%
  group_by(event) %>%
  shapiro_test(ryff)

outcomedata %>%
  group_by(event) %>%
  shapiro_test(pss)

outcomedata %>%
  group_by(event) %>%
  shapiro_test(bdi)

outcomedata %>%
  group_by(event) %>%
  shapiro_test(bai)

outcomedata %>%
  group_by(event) %>%
  shapiro_test(fmqi)

outcomedata %>%
  group_by(event) %>%
  shapiro_test(cerq)
```
Nonnormal: BDI @ baseline, post
           BAI @ baseline, post, followup
           MI @ followup
           CERQ @ post
           
### Reliability
```{r}
filename <- "20230330_Reliability.csv"
reliability <- read_csv(filename)
reliability <- reliability[1:104,]
```

```{r}
#ryff reliability
reliability_ryff <- reliability[,c("event", "rwbs_1", "rwbs_2", "rwbs_3", "rwbs_4", "rwbs_5", "rwbs_6", "rwbs_7", "rwbs_8","rwbs_9", "rwbs_10", "rwbs_11", "rwbs_12", "rwbs_13", "rwbs_14", "rwbs_15", "rwbs_16", "rwbs_17", "rwbs_18")]

reliability_ryff0 <- filter(reliability_ryff, event == "baseline")
reliability_ryff1 <- filter(reliability_ryff, event == "post")
reliability_ryff2 <- filter(reliability_ryff, event == "followup")

reliability_ryff0 <- reliability_ryff0[,-1]
reliability_ryff1 <- reliability_ryff1[,-1]
reliability_ryff2 <- reliability_ryff2[,-1]

#reverse 4, 7 , 8, 11, 12, 14, 17, 18
alpha(reliability_ryff0, keys = c(1, 1, 1, -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, -1, 1, 1, -1, -1))
alpha(reliability_ryff1,  keys = c(1, 1, 1, -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, -1, 1, 1, -1, -1))
alpha(reliability_ryff2,  keys = c(1, 1, 1, -1, 1, 1, -1, -1, 1, 1, -1, -1, 1, -1, 1, 1, -1, -1))
```
```{r}
#pss reliability
reliability_pss <- reliability[,c("event", "pss_1", "pss_2", "pss_3", "pss_4", "pss_5", "pss_6", "pss_7", "pss_8", "pss_9", "pss_10")]

reliability_pss0 <- filter(reliability_pss, event == "baseline")
reliability_pss1 <- filter(reliability_pss, event == "post")
reliability_pss2 <- filter(reliability_pss, event == "followup")

reliability_pss0 <- reliability_pss0[,-1]
reliability_pss1 <- reliability_pss1[,-1]
reliability_pss2 <- reliability_pss2[,-1]

#reverse 4, 5, 8 
alpha(reliability_pss0, keys = c(1, 1, 1, -1, -1, 1, 1, -1, 1, 1))
alpha(reliability_pss1,  keys = c(1, 1, 1, -1, -1, 1, 1, -1, 1, 1))
alpha(reliability_pss2,  keys = c(1, 1, 1, -1, -1, 1, 1, -1, 1, 1))
```
```{r}
#bdi reliability
reliability_bdi <- reliability[,c("event", "bdi_self1", "bdi_self2", "bdi_self3", "bdi_self4", "bdi_self5", "bdi_self6", "bdi_self7", "bdi_self8", "bdi_self10", "bdi_self11", "bdi_self12", "bdi_self13", "bdi_self14", "bdi_self15", "bdi_self16", "bdi_self17", "bdi_self18", "bdi_self19", "bdi_self20", "bdi_self21")]

reliability_bdi0 <- filter(reliability_bdi, event == "baseline")
reliability_bdi1 <- filter(reliability_bdi, event == "post")
reliability_bdi2 <- filter(reliability_bdi, event == "followup")

reliability_bdi0 <- reliability_bdi0[,-1]
reliability_bdi1 <- reliability_bdi1[,-1]
reliability_bdi2 <- reliability_bdi2[,-1]

alpha(reliability_bdi0)
alpha(reliability_bdi1)
alpha(reliability_bdi2)
```
```{r}
#bai reliability
reliability_bai <- reliability[,c("event", "beckself1", "beckself2", "beckself3", "beckself4", "beckself5", "beckself6", "beckself7", "beckself8", "beckself9", "beckself10", "beckself11", "beckself12", "beckself13", "beckself14", "beckself15", "beckself16", "beckself17", "beckself18", "beckself19", "beckself20", "beckself21")]

reliability_bai0 <- filter(reliability_bai, event == "baseline")
reliability_bai1 <- filter(reliability_bai, event == "post")
reliability_bai2 <- filter(reliability_bai, event == "followup")

reliability_bai0 <- reliability_bai0[,-1]
reliability_bai1 <- reliability_bai1[,-1]
reliability_bai2 <- reliability_bai2[,-1]

alpha(reliability_bai0)
alpha(reliability_bai1)
alpha(reliability_bai2)
```
```{r}
#mi reliability
reliability_mi <- reliability[,c("event", "fmi_q1", "fmi_q2", "fmi_q3", "fmi_q4", "fmi_q5", "fmi_q6", "fmi_q7", "fmi_q8", "fmi_q9", "fmi_q10", "fmi_q11", "fmi_q12", "fmi_q13", "fmi_q14")]

reliability_mi0 <- filter(reliability_mi, event == "baseline")
reliability_mi1 <- filter(reliability_mi, event == "post")
reliability_mi2 <- filter(reliability_mi, event == "followup")

reliability_mi0 <- reliability_mi0[,-1]
reliability_mi1 <- reliability_mi1[,-1]
reliability_mi2 <- reliability_mi2[,-1]

#reverse 13
alpha(reliability_mi0, keys = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, -1, 1))
alpha(reliability_mi1,  keys = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, -1, 1))
alpha(reliability_mi2,  keys = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, -1, 1))
```

```{r}
#cerq reliability
reliability_cerq <- reliability[,c("event","cerq_6_pos", "cerq_15_pos", "cerq_24_pos", "cerq_33_pos" )]

reliability_cerq0 <- filter(reliability_cerq, event == "baseline")
reliability_cerq1 <- filter(reliability_cerq, event == "post")
reliability_cerq2 <- filter(reliability_cerq, event == "followup")

reliability_cerq0 <- reliability_cerq0[,-1]
reliability_cerq1 <- reliability_cerq1[,-1]
reliability_cerq2 <- reliability_cerq2[,-1]

alpha(reliability_cerq0)
alpha(reliability_cerq1)
alpha(reliability_cerq2)
```

### Transformation
```{r}
outcomedata_t <- outcomedata

# box cox transformation of BDI
hist(outcomedata$bdi)
boxcox(lm(outcomedata$bdi ~ 1)) #0.5 - sqrt()
outcomedata_t$bdi <- sqrt(outcomedata$bdi)
hist(outcomedata_t$bdi)
```

```{r}
# boxcox transformation of BAI
hist(outcomedata$bai)
boxcox(lm(outcomedata$bdi ~ 1)) #0.5 - sqrt()
outcomedata_t$bai <- sqrt(outcomedata$bai)
hist(outcomedata_t$bai)
```
```{r}
# boxcox transformation of MI
hist(outcomedata$fmqi)
boxcox(lm(outcomedata$fmqi ~ 1)) #0.5 - sqrt()
outcomedata_t$fmqi <- sqrt(outcomedata$fmqi)
hist(outcomedata_t$fmqi)
```
```{r}
# boxcox transformation of CERQ
hist(outcomedata$cerq)
boxcox(lm(outcomedata$cerq ~ 1)) #0.5 - sqrt()
outcomedata_t$cerq <- sqrt(outcomedata$cerq)
hist(outcomedata_t$cerq)
```
```{r}
#recheck normality assumption
outcomedata_t %>%
  group_by(event) %>%
  shapiro_test(bdi)

outcomedata_t %>%
  group_by(event) %>%
  shapiro_test(bai)

outcomedata_t %>%
  group_by(event) %>%
  shapiro_test(fmqi)

outcomedata_t %>%
  group_by(event) %>%
  shapiro_test(cerq)
```

## Primary Analysis
### Mixed Models

```{r}
ryff_mm <- lmer(ryff ~ event * group + (1|id), data=outcomedata)
anova(ryff_mm)
summary(ryff_mm)


lme.dscore(ryff_mm, data=outcomedata, type="lme4") #calculate cohen's d for each effect in lme4 object
```

```{r}
pss_mm <- lmer(pss ~ event * group + (1|id), data=outcomedata)
anova(pss_mm)
summary(pss_mm)

lme.dscore(pss_mm, data=outcomedata, type="lme4")
```

```{r}
bdi_mm <- lmer(bdi ~ event * group + (1|id), data=outcomedata_t)
anova(bdi_mm)
summary(bdi_mm)

lme.dscore(bdi_mm, data=outcomedata, type="lme4")
```

```{r}
bai_mm <- lmer(bai ~ event * group + (1|id), data=outcomedata_t)

anova(bai_mm)
summary(bai_mm)

lme.dscore(bai_mm, data=outcomedata, type="lme4")
```

```{r}
mi_mm <- lmer(fmqi ~ event * group + (1|id), data=outcomedata_t)
anova(mi_mm)
summary(mi_mm)

lme.dscore(mi_mm, data=outcomedata, type="lme4")
```

```{r}
cerq_mm <- lmer(cerq ~event * group + (1|id), data=outcomedata_t)
anova(cerq_mm)
summary(cerq_mm)

lme.dscore(cerq_mm, data=outcomedata, type="lme4")
```

### Plots

```{r}
ggplot(outcomedata, aes(x = event, y = ryff)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("wellbeing") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

```{r}
ggplot(outcomedata, aes(x = event, y = pss)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("stress") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

```{r}
ggplot(outcomedata_t, aes(x = event, y = bdi)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("depression") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

```{r}
ggplot(outcomedata_t, aes(x = event, y = bai)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("anxiety") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

```{r}
ggplot(outcomedata_t, aes(x = event, y = fmqi)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("mindfulness") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

```{r}
ggplot(outcomedata_t, aes(x = event, y = cerq)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("cognitive emotion") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

# PRE-POST SESSION
```{r}
#load dataset
filename <- "20230324_PrePostData.csv"

prepostdata <- read_csv(filename)
```

```{r}
#reformat data: pre
predata <- prepostdata %>%
  dplyr::select(id, session, pos = pospre, neg = negpre, ios = iospre)

#create time variable
predata$time <- as.integer(0)
predata <- relocate(predata, time, .after=session)
```

```{r}
#reformat data: post
postdata <- prepostdata %>%
  dplyr::select(id, session, pos = pospost, neg = negpost, ios = iospost)

#create time variable
postdata$time <- as.integer(1)
postdata <- relocate(postdata, time, .after=session)
```

```{r}
#join predata and postdata
prepostdata <- merge(predata, postdata, all=TRUE)

prepostdata$id <- factor(prepostdata$id) 
prepostdata$session <- factor(prepostdata$session)
prepostdata$time <- factor(prepostdata$time)
levels(prepostdata$time) <- c("pre", "post")

view(prepostdata)
```

## Descriptive Statistics
### Mean (SD) Table
```{r}
# score means by time
prepostdata %>%
  group_by(time) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```

```{r}
# score means by session
prepostdata %>%
  group_by(session) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```
### Normality & Distribution
```{r}
# boxplot pos

box_pos <- ggplot(prepostdata, aes(x= session, y= pos, fill= time)) + scale_fill_manual(values = c("lightblue3", "dodgerblue4")) + geom_boxplot() + xlab("session") + ylab("score") + ggtitle("Positive Affect across Sessions")

box_pos
```

```{r}
# boxplot neg

box_neg <- ggplot(prepostdata, aes(x= session, y= neg, fill= time)) + scale_fill_manual(values = c("lightblue3", "dodgerblue4")) + geom_boxplot() + xlab("session") + ylab("score") + ggtitle("Negative Affect across Sessions")

box_neg
```
```{r}
# boxplot ios

box_ios <- ggplot(prepostdata, aes(x= session, y= ios, fill= time)) + scale_fill_manual(values = c("lightblue3", "dodgerblue4")) + geom_boxplot() + xlab("session") + ylab("score") + ggtitle("Connection to MT across Sessions")

box_ios
```
```{r}
# check for outliers
prepostdata %>%
  group_by(time) %>%
  identify_outliers(pos)

prepostdata %>%
  group_by(time) %>%
  identify_outliers(neg)

prepostdata %>%
  group_by(time) %>%
  identify_outliers(ios)
```
extreme outliers:
  12035 session 5, post
  12039 session 5, pre
  12039 session 6, pre
  12039 session 6, post

```{r}
#remove outliers
predata01 <- predata[-which(predata$id == "12039" & predata$session == "5"),]
predata01 <- predata01[-which(predata01$id == "12039" & predata01$session == "6"),]


predata01 %>%
  group_by(time) %>%
  identify_outliers(neg)
``` 

### Reliability

```{r}
# Calculate McDonald's Omega for PANAS
reliability <- read_csv("20230403_PrePostReliability.csv")
reliability_0 <- filter(reliability, time == "0")
reliability_1 <- filter(reliability, time == "1")

reliability_0 <- reliability_0[,4:23]
reliability_1 <- reliability_1[,4:23]
```

```{r}
describe(reliability_0)
omega(reliability_0)
```

```{r}
describe(reliability_1)
omega(reliability_1)
```


## Mixed Models
### Across Sessions 

WITH OUTLIERS
```{r}
#w/ outliers
lmm_pospre <- lmer(pos ~ session + (1|id), data=predata)

anova(lmm_pospre)
summary(lmm_pospre)

lme.dscore(lmm_pospre, data=predata, type="lme4")
```

```{r}
lmm_negpre <- lmer(neg ~ session + (1|id), data=predata)

anova(lmm_negpre)
summary(lmm_negpre)

lme.dscore(lmm_negpre, data=predata, type="lme4")
```

```{r}
lmm_iospre <- lmer(ios ~ session + (1|id), data=predata)

anova(lmm_iospre)
summary(lmm_iospre)

lme.dscore(lmm_iospre, data=predata, type="lme4")
```

WITHOUT OUTLIERS

```{r}
lmm_negpre01 <- lmer(neg ~ session + (1|id), data=predata01)

anova(lmm_negpre01)
summary(lmm_negpre01)

lme.dscore(lmm_negpre01, data=predata01, type="lme4")
```

#### Plots
```{r}
pos_across_sessions <- ggplot(data = predata, aes(x = session, y = pos)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("positive affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

pos_across_sessions
```

```{r}
neg_across_sessions <- ggplot(data = predata, aes(x = session, y = neg)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("negative affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

neg_across_sessions
```

```{r}
ios_across_sessions <- ggplot(data = predata, aes(x = session, y = ios)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("connection to therapist")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

ios_across_sessions
```
### Pre-post Session
```{r}
lmm_pos <- lmer(pos ~ time + (1|id), data=prepostdata)

anova(lmm_pos)
summary(lmm_pos)

lme.dscore(lmm_pos, data=prepostdata, type="lme4")
```

```{r}
lmm_neg <- lmer(neg ~ time + (1|id), data=prepostdata)

anova(lmm_neg)
summary(lmm_neg)

lme.dscore(lmm_neg, data=prepostdata, type="lme4")
```

```{r}
lmm_ios <- lmer(ios ~ time + (1|id), data=prepostdata)

anova(lmm_ios)
summary(lmm_ios)

lme.dscore(lmm_ios, data=prepostdata, type="lme4")
```
#### Plots
```{r}
pos_prepost <- ggplot(data = prepostdata, aes(x = time, y = pos)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("positive affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

pos_prepost
```

```{r}
neg_prepost <- ggplot(data = prepostdata, aes(x = time, y = neg)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("negative affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

neg_prepost
```

```{r}
ios_prepost <- ggplot(data = prepostdata, aes(x = time, y = ios)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("connection")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

ios_prepost
```
# EMA
```{r}
filename <- "20230321_EMAData.csv"

EMAdata <- read_csv(filename)
```

```{r}
#transform variables to factor
EMAdata$id <- as.factor(EMAdata$id)
EMAdata$event <- as.factor(EMAdata$event) 
levels(EMAdata$event) <- c("Baseline", "Session 4", "Post")
EMAdata$enrollmentgroup <- as.factor(EMAdata$enrollmentgroup) 
levels(EMAdata$enrollmentgroup) <- c("Waitlist Control", "Immediate Enrollment", "Group")
```

## Descriptive Statistics
```{r}
#remove group 2
EMAdata <- filter(EMAdata, enrollmentgroup != "Group")
```

### Missing Data

Missing Instances
```{r}
#create dataframe of missing instances (completed = 0)
missingdata <- filter(EMAdata, completed == '0')
```

```{r}
#create table of n missing: event x group
missing_instance <- matrix(0, 4,4)

colnames(missing_instance) <- c("Baseline", "Session4", "Post", "Total")
rownames(missing_instance) <- c("WLC", "IE", "Group", "Total")
```

```{r}
missing_instance[1,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "0"))
missing_instance[2,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "1"))
missing_instance[3,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "2"))
missing_instance[1,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "0"))
missing_instance[2,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "1"))
missing_instance[3,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "2"))
missing_instance[1,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "0"))
missing_instance[2,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "1"))
missing_instance[3,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "2"))

missing_instance[1,4] <- length(which(missingdata$enrollmentgroup == "0"))
missing_instance[2,4] <- length(which(missingdata$enrollmentgroup == "1"))
missing_instance[3,4] <- length(which(missingdata$enrollmentgroup == "2"))

missing_instance[4,1] <- length(which(missingdata$event == "baseline"))
missing_instance[4,2] <- length(which(missingdata$event == "session4"))
missing_instance[4,3] <- length(which(missingdata$event == "post"))
missing_instance[4,4] <- missing_instance[4,1] + missing_instance[4,2] + missing_instance[4,3]

view(missing_instance)
```

Missing Items
```{r}
#create dataframe of missing items (completed =1, N/A on at least 1 item)
missingitems <- filter(EMAdata, completed == "1" & (is.na(positive_score) | is.na(negative_score))) #n=9

view(missingitems)
```

```{r}
#impute missing items using the average of other neg/pos items
## create function

#function inputs: ID, EVENT, INSTANCE
itemavg <- function(ID, EVENT, INSTANCE){

  # row from id, event, instance
  r <- which(EMAdata$id == ID & EMAdata$event == EVENT & EMAdata$instance == INSTANCE)  
  
  positems <- c()
  negitems <- c()

  #positive items: 1, 3, 5, 9, 10, 12, 14, 16, 17, 19
  positems <- append(positems, as.integer(EMAdata[r,15])) #posq1
  positems <- append(positems, as.integer(EMAdata[r,17])) #posq3
  positems <- append(positems, as.integer(EMAdata[r,19]))#posq5
  positems <- append(positems, as.integer(EMAdata[r,23]))#posq9
  positems <- append(positems, as.integer(EMAdata[r,24])) #posq10
  positems <- append(positems, as.integer(EMAdata[r,26])) #posq12
  positems <- append(positems, as.integer(EMAdata[r,28])) #posq14
  positems <- append(positems, as.integer(EMAdata[r,30])) #posq16
  positems <- append(positems, as.integer(EMAdata[r,31])) #posq17
  positems <- append(positems, as.integer(EMAdata[r,33])) #posq19

  posavg <- mean(positems, na.rm=TRUE)

  #negative items: 2, 4, 6, 7, 8, 11, 13, 15, 18, 20
  negitems <- append(negitems,  as.integer(EMAdata[r,16])) #negq2
  negitems <- append(negitems, as.integer(EMAdata[r,18])) #negq4
  negitems <- append(negitems, as.integer(EMAdata[r,20])) #negq6
  negitems <- append(negitems, as.integer(EMAdata[r,21])) #negq7
  negitems <- append(negitems, as.integer(EMAdata[r,22])) #negq8
  negitems <- append(negitems, as.integer(EMAdata[r,25])) #negq11
  negitems <- append(negitems, as.integer(EMAdata[r,27])) #negq13
  negitems <- append(negitems, as.integer(EMAdata[r,29])) #negq15
  negitems <- append(negitems, as.integer(EMAdata[r,32])) #negq18
  negitems <- append(negitems, as.integer(EMAdata[r,34])) #negq20
  
  negavg <- mean(negitems, na.rm=TRUE)
  
  #output posavg, negavg
  out <- list(r, posavg, negavg)
  
  return(out)
}  
```

```{r}
#add all positive items
posscore <- function(r){
  score <- as.integer(EMAdata[r, 15]) + as.integer(EMAdata[r, 17]) + 
    as.integer(EMAdata[r, 19]) + as.integer(EMAdata[r, 23]) +
    as.integer(EMAdata[r, 24]) + as.integer(EMAdata[r, 26]) + 
    as.integer(EMAdata[r, 28]) + as.integer(EMAdata[r, 30]) + 
    as.integer(EMAdata[r, 31]) + as.integer(EMAdata[r, 33])
  
  return(score)
}

#add all negative items  
negscore <- function(r){
  score <- as.integer(EMAdata[r, 16]) + as.integer(EMAdata[r, 18]) + 
    as.integer(EMAdata[r, 20]) + as.integer(EMAdata[r, 21]) +
    as.integer(EMAdata[r, 22]) + as.integer(EMAdata[r, 25]) + 
    as.integer(EMAdata[r, 27]) + as.integer(EMAdata[r, 29]) + 
    as.integer(EMAdata[r, 32]) + as.integer(EMAdata[r, 34])
  
  return(score)
}
  
```

```{r}
#12015 session 4:8
out<- itemavg("12015", "Session 4", 8)
EMAdata[23, 17] <- out[2] #posq3 
EMAdata[23, 23] <- out[2] #posq9 
EMAdata[23,11] <- posscore(23) #positive score

#12015 session4: 12
out<- itemavg("12015", "Session 4", 12)
EMAdata[27, 25] <- out[3] #negq11
EMAdata[27,12] <- negscore(27) #negative score

#12017 post: 3
out <- itemavg("12017", "Post", 3)
EMAdata[123, 30] <- out[2] #posq16
EMAdata[123, 11] <- posscore(123)

#12027 baseline:2
out <- itemavg("12027", "Baseline", 2)
EMAdata[542, 28] <- out[2] #posq14
EMAdata[542, 11] <- posscore(542)

#12027 session4:12
out <- itemavg("12027", "Session 4", 12)
EMAdata[567, 22] <- out[3] #neg8
EMAdata[567, 27] <- out[3] #neg13
EMAdata[567,12] <- negscore(567)

#12033 baseline:15
out<- itemavg("12033", "Baseline", 15)
EMAdata[810, 24] <- out[2] #pos10
EMAdata[810, 11] <- posscore(810)

#12033 session4:11
out<- itemavg("12033", "Session 4", 11)
EMAdata[821, 30] <- out[2] #posq16
EMAdata[821, 11] <- posscore(821)

#12039 post:2
out<- itemavg("12039", "Post", 2)
EMAdata[1097, 31] <- out[2] #posq17
EMAdata[1097,11] <- posscore(1097)

#12039 post:3
out<- itemavg("12039", "Post", 3)
EMAdata[1098, 28] <- out[2] #posq14
EMAdata[1098, 34] <- out[3] #neg20
EMAdata[1098, 11] <- posscore(1098)
EMAdata[1098, 12] <- negscore(1098)
```

```{r}
#check missing items (completed =1, N/A on at least 1 item)
missingitems <- filter(EMAdata, completed == "1" & (is.na(positive_score) | is.na(negative_score))) #n=9

view(missingitems)
```

### Survey Completion

Compliance Rate

```{r}
#create dataframe of id x completion
compliance_rate <- matrix(0, 35, 3)

colnames(compliance_rate) <- c("Group", "Completed", "ComplianceRate")
id_names <- c("12015", "12016", "12017", "12018", "12019", "12020","12021", "12022", "12023", "12024","12025", "12026", "12027", "12028", "12029", "12030", "12031", "12032", "12033", "12034", "12035", "12036", "12037", "12038", "12039", "12040", "12041","12043", "12044", "12045", "12048","12050", "12055","12056", "12057")
rownames(compliance_rate) <- id_names
```

```{r}
#fill matrix col 1 with enrollment group
r = 1
n = 1

for(n in 1:35){
  ind <- which(EMAdata$id == id_names[n])
  compliance_rate[r,1] <- EMAdata$enrollmentgroup[ind[1]]
  r <- r + 1
  n <- n + 1
}

```

```{r}
#fill matrix col 2 with n completed by id
r = 1
n = 1

for(n in 1:35){
  compliance_rate[r,2] <- length(which(EMAdata$id == id_names[n]))
  r <- r + 1
  n <- n + 1
  }
```

```{r}
#fill matrix col 3 with compliance rate
r = 1
n = 1

for (n in 1:35){
  compliance_rate[r, 3] <- compliance_rate[r, 2]/45
  r <- r + 1
  n <- n + 1
}
```

```{r}
#output min and max values
max <- which(compliance_rate == max(compliance_rate[,3]), arr.ind=TRUE)
min <- which(compliance_rate == min(compliance_rate[,3]), arr.ind=TRUE)

print(paste("Maximum compliance rate:", compliance_rate[max[1]]))
print(paste("Minimum compliance rate:", compliance_rate[min]))
```

### Mean (SD) Table
```{r}
summarise_at(EMAdata, vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#group by event
EMAdata%>%
  group_by(event) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#group by enrollment
EMAdata%>%
  group_by(enrollmentgroup) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```
### Normality & Distribution
```{r}
#create grouped boxplot - positive

ggplot(EMAdata, aes(x=event, y=positive_score, fill=enrollmentgroup)) + scale_fill_manual(values = c("lightblue3", "dodgerblue4")) + geom_boxplot() + xlab("event") + ylab("score") + ggtitle("PANAS Positive Score x Group")
```

```{r}
#create grouped boxplot - negative

ggplot(EMAdata, aes(x=event, y=negative_score, fill=enrollmentgroup)) + 
  scale_fill_manual(values = c("lightblue3", "dodgerblue4")) +
  geom_boxplot() + xlab("event") + ylab("score") +
  ggtitle("PANAS Negative Score x Group")
```
## Reformat Variables
### Within-day

```{r}
withinday <- EMAdata[,c('id', 'event', 'day', 'instance', 'enrollmentgroup', 'time_interval', 'positive_score', 'negative_score', "positive_SD","negative_SD")]
```

```{r}
#convert time_interval from h:m:s to minutes
withinday$time_interval_min <- hour(withinday$time_interval)*60 + minute(withinday$time_interval)

withinday <- withinday %>%
  relocate("time_interval_min", .after="time_interval")
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Negative")
```

```{r}
#ASD = positive_SD[i] / time_interval[i]
pos_ASD <- vector()

for(i in 1:1427){
  pos_ASD <- append(pos_ASD, c(withinday$positive_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(pos_ASD)
```

```{r}
#create positive ASD variable
withinday$positive_ASD <- pos_ASD
```

```{r}
neg_ASD <- vector()

for(i in 1:1427){
  neg_ASD <- append(neg_ASD, c(withinday$negative_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(neg_ASD)
```

```{r}
#create negative ASD variable
withinday$negative_ASD <- neg_ASD
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Negative")
```
### Between-day
calculated in excel - kept running into issues doing it in R
```{r}
filename <- "20230325_betweenday.csv"

betweenday <- read_csv(filename)
```
## Preliminary Analyses
mixed model w REML
fixed effects: enrollment group, event, event*enrollmentgroup
random effects: id

### x Events
```{r}
lmm_emapos <- lmer(positive_mean ~ enrollmentgroup * event + (1|id), data=betweenday)

summary(lmm_emapos)
anova(lmm_emapos) 

lme.dscore(lmm_emapos, data=betweenday, type="lme4")
```

```{r}
lmm_emaneg <- lmer(negative_mean ~ enrollmentgroup * event + (1|id), data=betweenday)

summary(lmm_emaneg)
anova(lmm_emaneg) 

lme.dscore(lmm_emaneg, data=betweenday, type="lme4")
```

#### Plots
```{r}
# Positive Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panaspos <- confint(lmm_emapos, method="boot", nsim=10)
ci_panaspos <- tibble::rownames_to_column(data.frame(ci_panaspos), "Term") 
colnames(ci_panaspos)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panaspos)

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=positive_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Timepoints") + 
  labs(y = "positive affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lmm_emapos)[1], #Regression Line (RL).
              slope=fixef(lmm_emapos)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panaspos$`CI 97.5%`[3], 
              slope=ci_panaspos$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panaspos$`CI 2.5%`[3], 
              slope=ci_panaspos$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL
  
panaspos_ggplot
```

```{r}
# Negative Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panasneg <- confint(lmm_emaneg, method="boot", nsim=10)
ci_panasneg <- tibble::rownames_to_column(data.frame(ci_panasneg), "Term") 
colnames(ci_panasneg)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panasneg)

#Plot model 
panasneg_ggplot<- ggplot(data= betweenday, 
       aes(y=negative_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Timepoints") + 
  labs(y = "negative affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lmm_emaneg)[1], #Regression Line (RL).
              slope=fixef(lmm_emaneg)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panasneg$`CI 97.5%`[3], 
              slope=ci_panasneg$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panasneg$`CI 2.5%`[3], 
              slope=ci_panasneg$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL

panasneg_ggplot
```

### x Days
```{r}
betweenday$day <- as.factor(betweenday$day)
levels(betweenday$day) <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")
```

```{r}
lmm_emapos <- lmer(positive_mean ~ enrollmentgroup * day + (1|id), data=betweenday)

anova(lmm_emapos) 
summary(lmm_emapos)

lme.dscore(lmm_emapos, data=betweenday, type="lme4")
```

```{r}
lmm_emaneg <- lmer(negative_mean ~ enrollmentgroup * day + (1|id), data=betweenday)

anova(lmm_emaneg) 
summary(lmm_emaneg)

lme.dscore(lmm_emaneg, data=betweenday, type="lme4")
```

#### Plots
```{r}
# Positive Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=positive_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Days") + 
  labs(y = "positive affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) 
 
panaspos_ggplot
```

```{r}
# Negative Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday, 
       aes(y=negative_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Days") + 
  labs(y = "negative affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels)

panaspos_ggplot
```

### x Instances
```{r}
EMAdata$instance <- as.factor(EMAdata$instance)
levels(EMAdata$instance) <- c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15")
```

```{r}
lmm_emapos <- lmer(positive_score ~ enrollmentgroup * instance + (1|id), data= EMAdata)

anova(lmm_emapos) 
summary(lmm_emapos)

lme.dscore(lmm_emapos, data=EMAdata, type="lme4")
```

```{r}
lmm_emaneg <- lmer(negative_score ~ enrollmentgroup * instance + (1|id), data= EMAdata)

anova(lmm_emaneg) 
summary(lmm_emaneg)

lme.dscore(lmm_emaneg, data=EMAdata, type="lme4")
```

#### Plots
```{r}
# Positive Affect plot @ Baseline
EMA_baseline <- filter(EMAdata, event == "Baseline")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_baseline, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Baseline") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Positive Affect plot @ Session 4
EMA_session4 <- filter(EMAdata, event == "Session 4")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_session4, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Session4") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Positive Affect plot @ Post
EMA_post <- filter(EMAdata, event == "Post")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_post, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Post") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```

```{r}
# Negative Affect plot @ Baseline

instance_neg<- ggplot(data= EMA_baseline, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Baseline") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```

```{r}
# Negative Affect plot @ Session 4

instance_neg<- ggplot(data= EMA_session4, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Session 4") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```

```{r}
# Negative Affect plot @ Post

instance_neg<- ggplot(data= EMA_post, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Post") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```


