---
title: "20230705_MBMSPaper2_ZC"
output: html_document
date: "2023-07-05"
---

A combined document for all code used in Paper 2 analyses.

# Document Set Up
## Load Packages
```{r}
library(tidyverse) # data wrangling
library(dplyr) #data wrangling
library(lme4) # mixed models
library(ggpubr) # plots
library(readr) # read csv
library(psych) # psych statistics
library(rstatix) # outliers function
```

## Load Datasets
### a. Outcome Data
```{r}
#phase 2 outcome dataset
filename <- "20230621_phase2data.csv"

data <- read_csv(filename, col_types = cols(id = col_character()))
```

```{r}
data <- filter(data, group != "2") # remove Group

data$id <- factor(data$id) 
data$event <- factor(data$event, labels = c("baseline", "post", "followup"))
data$group <- factor(data$group, labels = c("control", "treatment"))

view(data)
```
### b. Pre-Post Session Data
```{r}
#read in data
prepostdata <- read_csv("20230324_PrePostData.csv")
```

```{r}
#reformat data: pre
predata <- prepostdata %>%
  select(id, session, pos = pospre, neg = negpre, ios = iospost)

#create time variable
predata$time <- as.integer(0)
predata <- relocate(predata, time, .after=session)
```

```{r}
#reformat data: post
postdata <- prepostdata %>%
  select(id, session, pos = pospost, neg = negpost, ios = iospost)

#create time variable
postdata$time <- as.integer(1)
postdata <- relocate(postdata, time, .after=session)
```

time 0 - pre
time 1 - post

```{r}
#join predata and postdata
prepostdata <- merge(predata, postdata, all=TRUE)

prepostdata$id <- factor(prepostdata$id) 
prepostdata$session <- factor(prepostdata$session)
prepostdata$time <- factor(prepostdata$time)
levels(prepostdata$time) <- c("pre", "post")

view(prepostdata)
```

### c. EMA Data
```{r}
#ema dataset

```

# Outcome Data
## 1. Descriptive Statistics

```{r}
data_baseline <- filter(data, event == "baseline")

data_baseline %>%
  group_by(group)%>%
  get_summary_stats(type = "mean_sd")

data_baseline%>%
  get_summary_stats(type = "mean_sd")
```

```{r}
data_baseline <- filter(data, event == "post")

data_baseline %>%
  group_by(group)%>%
  get_summary_stats(type = "mean_sd")

data_baseline%>%
  get_summary_stats(type = "mean_sd")
```

```{r}
data_baseline <- filter(data, event == "followup")

data_baseline %>%
  group_by(group)%>%
  get_summary_stats(type = "mean_sd")

data_baseline%>%
  get_summary_stats(type = "mean_sd")
```

```{r}
#BDI clinical significance (score >= 20)
bdi_wlc_sig <- which(data_baseline$bdi >=20 & data_baseline$group == 0) #wlc group
bdi_wlc_n <- length(bdi_wlc_sig)
bdi_wlc_perc <- (bdi_wlc_n/16) * 100
bdi_wlc_perc

bdi_ie_sig <- which(data_baseline$bdi >=20 & data_baseline$group == 1) #ie group
bdi_ie_n <- length(bdi_ie_sig)
bdi_ie_perc <- (bdi_ie_n/17) * 100
bdi_ie_perc

((bdi_wlc_n + bdi_ie_n)/33)*100 #total

#BAI clinical significance (score >= 16)
bai_wlc_sig <- which(data_baseline$bai >=16 & data_baseline$group == 0) #wlc group
bai_wlc_n <- length(bai_wlc_sig)
bai_wlc_perc <- (bai_wlc_n/16) * 100
bai_wlc_perc

bai_ie_sig <- which(data_baseline$bai >=16 & data_baseline$group == 1) #ie group
bai_ie_n <- length(bai_ie_sig)
bai_ie_perc <- (bai_ie_n/17) * 100
bai_ie_perc

((bai_wlc_n + bai_ie_n)/33)*100 #total
```

### a. Normality & Distribution
```{r}
#boxplots
bxp_ryff <- ggboxplot(data, x = "event", y = "ryff", add = "point")
bxp_ryff

bxp_pss <- ggboxplot(data, x = "event", y = "pss", add = "point")
bxp_pss

bxp_bdi <- ggboxplot(data, x = "event", y = "bdi", add = "point")
bxp_bdi

bxp_bai <- ggboxplot(data, x = "event", y = "bai", add = "point")
bxp_bai

bxp_mi <- ggboxplot(data, x = "event", y = "fmqi", add = "point")
bxp_mi

bxp_cerq <- ggboxplot(data, x = "event", y = "cerq", add = "point")
bxp_cerq
```

```{r}
#qq plots
qq_ryff <- ggqqplot(data, "ryff", facet.by = "event")
qq_ryff
qq_pss <- ggqqplot(data, "pss", facet.by = "event")
qq_pss
qq_bdi <- ggqqplot(data, "bdi", facet.by = "event")
qq_bdi
qq_bai <- ggqqplot(data, "bai", facet.by = "event")
qq_bai
qq_mi <- ggqqplot(data, "fmqi", facet.by = "event")
qq_mi
qq_cerq <- ggqqplot(data, "cerq", facet.by = "event")
qq_cerq
```

```{r}
#check for outliers
data %>%
  group_by(event) %>%
  identify_outliers(ryff)

data %>%
  group_by(event) %>%
  identify_outliers(pss)

data %>%
  group_by(event) %>%
  identify_outliers(bdi)

data %>%
  group_by(event) %>%
  identify_outliers(bai)

data %>%
  group_by(event) %>%
  identify_outliers(fmqi)

data %>%
  group_by(event) %>%
  identify_outliers(cerq)

```
no extreme outliers found

```{r}
# test for normality with shapiro wilks test
data %>%
  group_by(event) %>%
  shapiro_test(ryff)

data %>%
  group_by(event) %>%
  shapiro_test(pss)

data %>%
  group_by(event) %>%
  shapiro_test(bdi)

data %>%
  group_by(event) %>%
  shapiro_test(bai)

data %>%
  group_by(event) %>%
  shapiro_test(fmqi)

data %>%
  group_by(event) %>%
  shapiro_test(cerq)
```
Nonnormal: 
BDI @ baseline, post
BAI @ baseline, post, followup
MI @ followup
CERQ @ post

### b. Transformations
```{r}
#BDI transformation
hist(data$bdi, main="Original")
data$bdi_t <- sqrt(data$bdi)
hist(data$bdi_t, main="SQRT Transformed")
shapiro.test(data$bdi_t)
```
```{r}
#BAI transformation
hist(data$bai, main="Original")
data$bai_t <- sqrt(data$bai)
hist(data$bai_t, main="SQRT Transformed")
shapiro.test(data$bai_t)
```
```{r}
#MI transformation
hist(data$fmqi, main="Original")
data$fmqi_t <- sqrt(data$fmqi)
hist(data$fmqi_t, main="SQRT Transformed")
shapiro.test(data$fmqi_t)
```
```{r}
#CERQ transformation
hist(data$cerq, main="Original")
data$cerq_t <- sqrt(data$cerq)
hist(data$cerq_t, main="SQRT Transformed")
shapiro.test(data$cerq_t)
```

### c. Figures
Line graphs of non-transformed outcome data
```{r}
ggplot(data, aes(x = event, y = ryff)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Well-being") + xlab("") + #bold + #Add Y and X axis 
  theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=22)) + theme(axis.title.x = element_text(size=22)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 18)) + theme(legend.text = element_text(size=18)) #Increase font size of legend
```

```{r}
ggplot(data, aes(x = event, y = pss)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Perceived Stress") + xlab("Timepoint") + #Add Y and X axis
  theme(axis.text.x = element_text(size=16)) + theme(axis.text.y = element_text(size=16)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=20)) + theme(axis.title.x = element_text(size=20)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 16)) + theme(legend.text = element_text(size=14)) #Increase font size of legend
```

```{r}
ggplot(data, aes(x = event, y = bdi)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Depression") + xlab("Timepoint") + #Add Y and X axis
  theme(axis.text.x = element_text(size=16)) + theme(axis.text.y = element_text(size=16)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=20)) + theme(axis.title.x = element_text(size=20)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 16)) + theme(legend.text = element_text(size=14)) #Increase font size of legend
```

```{r}
ggplot(data, aes(x = event, y = bai)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Anxiety") + xlab("Timepoint") + #Add Y and X axis
  theme(axis.text.x = element_text(size=16)) + theme(axis.text.y = element_text(size=16)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=20)) + theme(axis.title.x = element_text(size=20)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 16)) + theme(legend.text = element_text(size=14)) #Increase font size of legend
```

```{r}
ggplot(data, aes(x = event, y = fmqi)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Mindfulness") + xlab("Timepoint") + #Add Y and X axis
  theme(axis.text.x = element_text(size=16)) + theme(axis.text.y = element_text(size=16)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=20)) + theme(axis.title.x = element_text(size=20)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 16)) + theme(legend.text = element_text(size=14)) #Increase font size of legend
```

```{r}
ggplot(data, aes(x = event, y = cerq)) + 
  theme_classic() + # Set white background, no lines
  stat_summary(aes(group=group, color=group, linetype=group), geom="line", size=3, fun.y = mean) + #Add mean line by Group
  scale_color_manual(values = c("grey70","grey4")) + #Customize colors
  # add confidence intervals for each group mean line?
  ylab("Cognitive Reappraisal") + xlab("Timepoint") + #Add Y and X axis
  theme(axis.text.x = element_text(size=16)) + theme(axis.text.y = element_text(size=16)) + #Increase font size of axes
  theme(axis.title.y = element_text(size=20)) + theme(axis.title.x = element_text(size=20)) + #Increase font size of axis titles
  theme(legend.title = element_text(size = 16)) + theme(legend.text = element_text(size=14)) #Increase font size of legend
```

## 2. Primary Analysis
```{r}
ryff_mm <- lmer(ryff ~ event * group + (1|id), data=data)
anova(ryff_mm)
summary(ryff_mm)
```

```{r}
pss_mm <- lmer(pss ~ event * group + (1|id), data=data)
anova(pss_mm)
summary(pss_mm)
```

```{r}
bdi_mm <- lmer(bdi ~ event + group + event * group + (1|id), data=data)
anova(bdi_mm)
summary(bdi_mm)
```

```{r}
bai_mm <- lmer(bai ~ event + group + event * group + (1|id), data=data)
anova(bai_mm)
summary(bai_mm)
```

```{r}
mi_mm <- lmer(fmqi ~ event + group + event * group + (1|id), data=data)
anova(mi_mm)
summary(mi_mm)
```

```{r}
cerq_mm <- lmer(cerq ~ event + group + event * group + (1|id), data=data)
anova(cerq_mm)
summary(cerq_mm)
```

# Pre-Post Session Data
## 1. Descriptive Statistics
```{r}
# score means by time
prepostdata %>%
  group_by(time) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```

```{r}
# score means by session
prepostdata %>%
  group_by(session) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```

### a. Normality & Distribution
```{r}
# boxplot
ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$pos, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Positive Affect across Sessions")

ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$neg, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Negative Affect across Sessions")

ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$ios, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Connection to MT across Sessions")
```

```{r}

```

# EMA Data