---
title: "20220117_BaselinePostFollowup_Phase2"
output: html_document
---
### Load Packages
```{r}
library(tidyverse) # data wrangling
library(dplyr) #data wrangling
library(lme4) # mixed models
library(ggpubr) # plots
library(readr) # read csv
```

### Load Dataset
```{r}
filename <- "20230313_phase2data.csv"

data <- read_csv(filename, col_types = cols(id = col_character()))
```

```{r}
data <- filter(data, group != "2") # remove Group

data$id <- factor(data$id) 
data$event <- factor(data$event, labels = c("baseline", "post", "followup"))
data$group <- factor(data$group, labels = c("control", "treatment"))

view(data)
```

### Coded Variables
**Event**
1 - Baseline
2 - Post
3 - Followup

**Group**
0 - Waitlist Control
1 - MBMS Enrollment
2 - Group

# Descriptive Statistics #

```{r}
library(psych)
describe(data[,4:9])
```

```{r}
data_baseline <- filter(data, event == "baseline")

data_baseline %>%
  group_by(group)%>%
  get_summary_stats(type = "mean_sd")

data_baseline%>%
  get_summary_stats(type = "mean_sd")
```

```{r}
#BDI clinical significance (score >= 20)
bdi_wlc_sig <- which(data_baseline$bdi >=20 & data_baseline$group == 0) #wlc group
bdi_wlc_n <- length(bdi_wlc_sig)
bdi_wlc_perc <- (bdi_wlc_n/16) * 100
bdi_wlc_perc

bdi_ie_sig <- which(data_baseline$bdi >=20 & data_baseline$group == 1) #ie group
bdi_ie_n <- length(bdi_ie_sig)
bdi_ie_perc <- (bdi_ie_n/17) * 100
bdi_ie_perc

((bdi_wlc_n + bdi_ie_n)/33)*100 #total

#BAI clinical significance (score >= 16)
bai_wlc_sig <- which(data_baseline$bai >=16 & data_baseline$group == 0) #wlc group
bai_wlc_n <- length(bai_wlc_sig)
bai_wlc_perc <- (bai_wlc_n/16) * 100
bai_wlc_perc

bai_ie_sig <- which(data_baseline$bai >=16 & data_baseline$group == 1) #ie group
bai_ie_n <- length(bai_ie_sig)
bai_ie_perc <- (bai_ie_n/17) * 100
bai_ie_perc

((bai_wlc_n + bai_ie_n)/33)*100 #total
```

## Normality & Distributions
### Create Boxplot for each variable
```{r}
bxp_ryff <- ggboxplot(data, x = "event", y = "ryff", add = "point")
bxp_ryff

bxp_pss <- ggboxplot(data, x = "event", y = "pss", add = "point")
bxp_pss

bxp_bdi <- ggboxplot(data, x = "event", y = "bdi", add = "point")
bxp_bdi

bxp_bai <- ggboxplot(data, x = "event", y = "bai", add = "point")
bxp_bai

bxp_mi <- ggboxplot(data, x = "event", y = "fmqi", add = "point")
bxp_mi

bxp_cerq <- ggboxplot(data, x = "event", y = "cerq", add = "point")
bxp_cerq
```

###Create QQ plot for each variable
```{r}
qq_ryff <- ggqqplot(data, "ryff", facet.by = "event")
qq_ryff
qq_pss <- ggqqplot(data, "pss", facet.by = "event")
qq_pss
qq_bdi <- ggqqplot(data, "bdi", facet.by = "event")
qq_bdi
qq_bai <- ggqqplot(data, "bai", facet.by = "event")
qq_bai
qq_mi <- ggqqplot(data, "fmqi", facet.by = "event")
qq_mi
qq_cerq <- ggqqplot(data, "cerq", facet.by = "event")
qq_cerq
```
###Check for outliers
```{r}
library(rstatix)

data %>%
  group_by(event) %>%
  identify_outliers(ryff)

data %>%
  group_by(event) %>%
  identify_outliers(pss)

data %>%
  group_by(event) %>%
  identify_outliers(bdi)

data %>%
  group_by(event) %>%
  identify_outliers(bai)

data %>%
  group_by(event) %>%
  identify_outliers(fmqi)

data %>%
  group_by(event) %>%
  identify_outliers(cerq)

```
no extreme outliers found

###Check normality Assumption
```{r}
data %>%
  group_by(event) %>%
  shapiro_test(ryff)

data %>%
  group_by(event) %>%
  shapiro_test(pss)

data %>%
  group_by(event) %>%
  shapiro_test(bdi)

data %>%
  group_by(event) %>%
  shapiro_test(bai)

data %>%
  group_by(event) %>%
  shapiro_test(fmqi)

data %>%
  group_by(event) %>%
  shapiro_test(cerq)
```
Nonnormal: BDI @ baseline, post
           BAI @ baseline, post, followup
           MI @ followup
           CERQ @ post
          
# Primary Analysis #


```{r}
# run without 12032 and 12046 (not missing at random) 
data_01 <- filter(data, id != 12032)
data_01 <- filter(data, id != 12046)
```

```{r}
ryff_mm <- lmer(ryff ~ event * group + (1|id), data=data)
anova(ryff_mm)
summary(ryff_mm)
```

```{r}
ryff_mm <- lmer(ryff ~ event * group + (1|id), data=data_01)
anova(ryff_mm)
summary(ryff_mm)
```

```{r}
pss_mm <- lmer(pss ~ event * group + (1|id), data=data)
anova(pss_mm)
summary(pss_mm)
```

```{r}
pss_mm <- lmer(pss ~ event * group + (1|id), data=data_01)
anova(pss_mm)
summary(pss_mm)
```

```{r}
bdi_mm <- lmer(bdi ~ event + group + event * group + (1|id), data=data)
anova(bdi_mm)
summary(bdi_mm)
```

```{r}
bdi_mm <- lmer(bdi ~ event + group + event * group + (1|id), data=data_01)
anova(bdi_mm)
summary(bdi_mm)
```

```{r}
bai_mm <- lmer(bai ~ event + group + event * group + (1|id), data=data)
anova(bai_mm)
summary(bai_mm)
```

```{r}
bai_mm <- lmer(bai ~ event + group + event * group + (1|id), data=data_01)
anova(bai_mm)
summary(bai_mm)
```

```{r}
mi_mm <- lmer(fmqi ~ event + group + event * group + (1|id), data=data)
anova(mi_mm)
summary(mi_mm)
```

```{r}
mi_mm <- lmer(fmqi ~ event + group + event * group + (1|id), data=data_01)
anova(mi_mm)
summary(mi_mm)
```

```{r}
cerq_mm <- lmer(cerq ~ event + group + event * group + (1|id), data=data)
anova(cerq_mm)
summary(cerq_mm)
```

```{r}
cerq_mm <- lmer(cerq ~ event + group + event * group + (1|id), data=data_01)
anova(cerq_mm)
summary(cerq_mm)
```

## MLM Plots

```{r}
ggplot(data, aes(x = event, y = ryff)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("wellbeing") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```
```{r}
ggplot(data, aes(x = event, y = pss)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("stress") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```
```{r}
ggplot(data, aes(x = event, y = bdi)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("depression") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```
```{r}
ggplot(data, aes(x = event, y = bai)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("anxiety") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```
```{r}
ggplot(data, aes(x = event, y = fmqi)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("mindfulness") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```
```{r}
ggplot(data, aes(x = event, y = cerq)) + 
  stat_summary(aes(group=group, color=group), geom="line", size=2, fun.y = mean) +
  ylab("cognitive emotion") + xlab("") + scale_color_manual(values = c("lightblue3","darkblue"))
```

# Pre-Post Session
```{r}
#read in data
prepostdata <- read_csv("20230324_PrePostData.csv")
```

```{r}
#reformat data: pre
predata <- prepostdata %>%
  select(id, session, pos = pospre, neg = negpre, ios = iospost)

#create time variable
predata$time <- as.integer(0)
predata <- relocate(predata, time, .after=session)
```

```{r}
#reformat data: post
postdata <- prepostdata %>%
  select(id, session, pos = pospost, neg = negpost, ios = iospost)

#create time variable
postdata$time <- as.integer(1)
postdata <- relocate(postdata, time, .after=session)
```

time 0 - pre
time 1 - post

```{r}
#join predata and postdata
prepostdata <- merge(predata, postdata, all=TRUE)

prepostdata$id <- factor(prepostdata$id) 
prepostdata$session <- factor(prepostdata$session)
prepostdata$time <- factor(prepostdata$time)
levels(prepostdata$time) <- c("pre", "post")

view(prepostdata)
```

## Descriptive Statistics
```{r}
# score means by time
prepostdata %>%
  group_by(time) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```
# normality
```{r}
# score means by session
prepostdata %>%
  group_by(session) %>%
  get_summary_stats(pos, neg, ios, type="mean_sd")
```

```{r}
# boxplot pos

box_pos <- ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$pos, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Positive Affect across Sessions")

box_pos
```
```{r}
# boxplot neg

box_neg <- ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$neg, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Negative Affect across Sessions")

box_neg
```

```{r}
# boxplot ios

box_ios <- ggplot(prepostdata, aes(x= prepostdata$session, y= prepostdata$ios, fill= prepostdata$time)) + 
  geom_boxplot() + xlab("Session") + ylab("Score") +
  ggtitle("Connection to MT across Sessions")

box_ios
```

```{r}
# outliers
prepostdata %>%
  group_by(time) %>%
  identify_outliers(pos)

prepostdata %>%
  group_by(time) %>%
  identify_outliers(neg)

  # 12035 session 5, post
  # 12039 session 5, pre
  # 12039 session 6, pre
  # 12039 session 6, post

prepostdata %>%
  group_by(time) %>%
  identify_outliers(ios)

```

## Linear Mixed Models
### session * time

```{r}
lmm_neg <- lmer(neg ~ session*time + (1|id), data=prepostdata)

anova(lmm_neg)
summary(lmm_neg)
```

```{r}
lmm_ios <- lmer(ios ~ session*time + (1|id), data=prepostdata)

anova(lmm_ios)
summary(lmm_ios)
```

#### plots ####
```{r}
pos_prepost <- ggplot(data = prepostdata, aes(x = time, y = pos)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("positive affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

pos_prepost
```

```{r}
neg_prepost <- ggplot(data = prepostdata, aes(x = time, y = neg)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("negative affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

neg_prepost
```
```{r}
ios_prepost <- ggplot(data = prepostdata, aes(x = time, y = ios)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("connection")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =1, group = 1)

ios_prepost
```
### across session only pre

```{r}
#w/ outliers
lmm_pospre <- lmer(pos ~ session + (1|id), data=predata)

anova(lmm_pospre)
summary(lmm_pospre)
```
```{r}
lmm_negpre <- lmer(neg ~ session + (1|id), data=predata)

anova(lmm_negpre)
summary(lmm_negpre)
```
```{r}
lmm_iospre <- lmer(ios ~ session + (1|id), data=predata)

anova(lmm_iospre)
summary(lmm_iospre)
```

#### plots ####
```{r}
pos_across_sessions <- ggplot(data = predata, aes(x = session, y = pos)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("positive affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

pos_across_sessions
```

```{r}
neg_across_sessions <- ggplot(data = predata, aes(x = session, y = neg)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("negative affect")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

neg_across_sessions
```

```{r}
ios_across_sessions <- ggplot(data = predata, aes(x = session, y = ios)) + geom_line(aes(group=id), size=0.5, color="lightblue3") + ylab("connection to therapist")  +theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size=20, face="bold")) + geom_line(stat="summary", color ="darkblue", size =2)

ios_across_sessions
```

```{r}
#remove outliers
predata01 <- predata[-which(predata$id == "12039" & predata$session == "5"),]
predata01 <- predata01[-which(predata01$id == "12039" & predata01$session == "6"),]


predata01 %>%
  group_by(time) %>%
  identify_outliers(neg)
```
```{r}
lmm_negpre01 <- lmer(neg ~ session + (1|id), data=predata01)

anova(lmm_negpre01)
summary(lmm_negpre01)
```

