---
title: "20220421_ema_dataanalysis"
output: pdf_document
---
last updated: 6/10/2022 by ZC

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(EMAtools)
```

import data set
```{r}
ema_df <- as_tibble(ema_data)
ema_df
```
create day of week variable
```{r}
ema_df$dayofweek <- weekdays(ema_df$date)

ema_df$cat_dayofweek <- ema_df$dayofweek
ema_df$cat_dayofweek <- recode(ema_df$cat_dayofweek, 
                               "Monday" = "Weekday", "Tuesday" = "Weekday",
                               "Wednesday" = "Weekday", "Thursday" = "Weekday",
                               "Friday" = "Weekday", "Saturday" = "Weekend", 
                               "Sunday" = "Weekend")

head(ema_df)
```

***Frequency Distributions***
Frequency distributions of timestamp by all IDs, by individual, and by event

```{r}
ema_df %>%
  mutate(hour = hour(time_participant)) %>%
  ggplot(aes(x=hour)) + geom_bar() + 
  ggtitle("Timestampe Freq Dist - All IDs")
```
```{r}
ema_df %>%
  mutate(hour = hour(time_participant)) %>%
  ggplot(aes(x=hour)) + geom_bar() + 
  ggtitle("Timestampe Freq Dist - by ID") +
  facet_wrap(~id)
```
graph by event
```{r}
ema_df %>%
  mutate(hour = hour(time_participant)) %>%
  ggplot(aes(x=hour)) + geom_bar() + 
  ggtitle("Timestampe Freq Dist - by event") +
  facet_wrap(~event, ncol=1)
```
***Survey Completion***
Calculation of number of surveys completed by participants.

```{r}
ema_df$completed <- as.integer(if_else(
  !is.na(ema_df$panaspos_score), 1, 0))

ema_df$instance <- as.integer(ema_df$instance)

head(ema_df)

```
create data frame for completion sums
```{r}
ema_completion <- data.frame(matrix(ncol=5, nrow=28))
colnames(ema_completion) <- c("id", "event", "group", "day", "completed")
```

populate ema_completion data frame and sum completed
```{r}
nrows = 28*3
ystart = 1
yend = 15

for(x in 1:nrows){
  count <- sum(ema_df[ystart:yend, 'completed'])
  
  ema_completion[x,1] <- ema_df[yend,1]
  ema_completion[x,2] <- ema_df[yend,2]
  ema_completion[x,3] <- ema_df[yend,4]
  ema_completion[x,4] <- ema_df[yend, 'cat_dayofweek']
  ema_completion[x,5] <- count
  
  ystart = ystart + 15
  yend = yend + 15
  x = x + 1
}
  
head(ema_completion)

```
```{r}
#mean completed for all IDs
surveys_mean_allids <- mean(ema_completion$completed)
surveys_sd_allids <- sd(ema_completion$completed)

surveys_mean_allids 
surveys_sd_allids
```

```{r}
#mean completed by event
ema_completion %>%
  group_by(event)%>%
  summarise_at(vars(completed), list(mean=mean, sd=sd))
```
```{r}
#mean completed by group
ema_completion %>%
  group_by(group)%>%
  summarise_at(vars(completed), list(mean=mean, sd=sd))
```
```{r}
#mean completed by ID
ema_completion %>%
  group_by(id)%>%
  summarise_at(vars(completed), list(mean=mean, sd=sd))
```
```{r}
#mean completed by day of week
ema_completion %>%
  group_by(day)%>%
  summarise_at(vars(completed), list(mean=mean, sd=sd))
```

***PANAS Score Descriptive***
```{r}
#mean score for all IDs
panaspos_mean_allids <- mean(ema_df$panaspos_score, na.rm=TRUE)
panaspos_sd_allids <- sd(ema_df$panaspos_score, na.rm=TRUE)

panasneg_mean_allids <- mean(ema_df$panasneg_score, na.rm=TRUE)
panasneg_sd_allids <- sd(ema_df$panasneg_score, na.rm=TRUE)

panaspos_mean_allids
panaspos_sd_allids
panasneg_mean_allids
panasneg_sd_allids
```

```{r}
#panas score mean by event
ema_df %>%
  group_by(event)%>%
  summarise_at(vars(panaspos_score, panasneg_score), 
               list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#panas score mean by group
ema_df %>%
  group_by(group)%>%
  summarise_at(vars(panaspos_score, panasneg_score), 
               list(mean=mean, sd=sd), na.rm=TRUE)
```

```{r}
#panasscore mean by ID
ema_df %>%
  group_by(id)%>%
  summarise_at(vars(panaspos_score, panasneg_score), 
               list(mean=mean, sd=sd), na.rm=TRUE)
```
```{r}
#panas score mean by day
ema_df %>%
  group_by(cat_dayofweek)%>%
  summarise_at(vars(panaspos_score, panasneg_score), 
               list(mean=mean, sd=sd), na.rm=TRUE)
```

*****EMA MODEL*****
```{r}
library(modelr)
```

create variable - within-day successive difference
```{r}
model_df <- ema_df
model_df$date_time <- as.POSIXct(paste(model_df$date, model_df$time), 
format= "%Y-%m-%d %H:%M:%S")

model_df <- model_df[, colnames(model_df)[c(1:7, 35, 32, 33, 8:31, 34)]]
```

nrows = 28*3
ystart = 1
yend = 15

for(x in 1:nrows){
  count <- sum(ema_df[ystart:yend, 'completed'])
  
  ema_completion[x,1] <- ema_df[yend,1]
  ema_completion[x,2] <- ema_df[yend,2]
  ema_completion[x,3] <- ema_df[yend,4]
  ema_completion[x,4] <- ema_df[yend, 'cat_dayofweek']
  ema_completion[x,5] <- count
  
  ystart = ystart + 15
  yend = yend + 15
  x = x + 1
}
  
head(ema_completion)
```{r}
wd_sd <- c()
wd_sd <- c(wd_sd, 0)

x=2

for(x in 1:(nrows)){
  diff = ifelse(model_df$id[x] != model_df$id[x-1], 0,
              ifelse(model_df$date[x] != model_df$date[x-1], 0,
                (difftime(model_df$time[x], model_df$time[x-1], units = "mins"))))
  
  diff <- as.numeric(diff)

  wd_sd <- c(wd_sd, diff)
  x = x + 1
}
```

```{r}
length(wd_sd)

model_df$wd_sd <- wd_sd
model_df$wd_sd[wd_sd == 0] <- NA

model_df <- model_df[, colnames(model_df)[c(1:12, 36, 13:35)]]
```

