---
title: "20230210_EMAAnalysis"
output: html_document
date: "2023-02-10"
---

```{r setup, include=FALSE}
library(tidyverse) #data wrangling
library(dplyr) # data wrangling
library(ggpubr) # plotting
library(readr) # read_csv
library(psych) # psych statistics - MSSD
library(lme4) # mixed models
```

# Import Data

```{r}
EMAdata <- read_csv("20230214_EMAData.csv")
View(EMAdata)
```

```{r}
#transform variables to factor
EMAdata$id <- as.factor(EMAdata$id)
EMAdata$event <- as.factor(EMAdata$event) 
levels(EMAdata$event) <- c("Baseline", "Session 4", "Post")
EMAdata$enrollmentgroup <- as.factor(EMAdata$enrollmentgroup) 
levels(EMAdata$enrollmentgroup) <- c("Waitlist Control", "Immediate Enrollment", "Group")
```

# Descriptive Statistics
## Missing Data
### Missing Instances

```{r}
#create dataframe of missing instances (completed = 0)
missingdata <- filter(EMAdata, completed == '0')
```

```{r}
#create table of n missing: event x group
missing_instance <- matrix(0, 4,4)

colnames(missing_instance) <- c("Baseline", "Session4", "Post", "Total")
rownames(missing_instance) <- c("WLC", "IE", "Group", "Total")
```

```{r}
missing_instance[1,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "0"))
missing_instance[2,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "1"))
missing_instance[3,1] <- length(which(missingdata$event == "baseline" & missingdata$enrollmentgroup == "2"))
missing_instance[1,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "0"))
missing_instance[2,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "1"))
missing_instance[3,2] <- length(which(missingdata$event == "session4" & missingdata$enrollmentgroup == "2"))
missing_instance[1,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "0"))
missing_instance[2,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "1"))
missing_instance[3,3] <- length(which(missingdata$event == "post" & missingdata$enrollmentgroup == "2"))

missing_instance[1,4] <- length(which(missingdata$enrollmentgroup == "0"))
missing_instance[2,4] <- length(which(missingdata$enrollmentgroup == "1"))
missing_instance[3,4] <- length(which(missingdata$enrollmentgroup == "2"))

missing_instance[4,1] <- length(which(missingdata$event == "baseline"))
missing_instance[4,2] <- length(which(missingdata$event == "session4"))
missing_instance[4,3] <- length(which(missingdata$event == "post"))
missing_instance[4,4] <- missing_instance[4,1] + missing_instance[4,2] + missing_instance[4,3]
```

### Missing Items 

```{r}
#create dataframe of missing items (completed =1, N/A on at least 1 item)
missingitems <- filter(EMAdata, completed == "1" & (is.na(positive_score) | is.na(negative_score))) #n=9

#decision - listwise deletion (bc of low n of instances with missing items)
```

```{r}
nrow(EMAdata)

EMAdata <- filter(EMAdata, !is.na(positive_score))
EMAdata <- filter(EMAdata, !is.na(negative_score))

nrow(EMAdata)
```

## Survey Completion
### Comliance Rate

```{r}
#create dataframe of id x completion
compliance_rate <- matrix(0, 35, 3)

colnames(compliance_rate) <- c("Group", "Completed", "ComplianceRate")
id_names <- c("12015", "12016", "12017", "12018", "12019", "12020","12021", "12022", "12023", "12024","12025", "12026", "12027", "12028", "12029", "12030", "12031", "12032", "12033", "12034", "12035", "12036", "12037", "12038", "12039", "12040", "12041","12043", "12044", "12045", "12048","12050", "12055","12056", "12057")
rownames(compliance_rate) <- id_names
```

```{r}
#fill matrix col 1 with enrollment group
r = 1
n = 1

for(n in 1:35){
  ind <- which(EMAdata$id == id_names[n])
  compliance_rate[r,1] <- EMAdata$enrollmentgroup[ind[1]]
  r <- r + 1
  n <- n + 1
}

```

```{r}
#fill matrix col 2 with n completed by id
r = 1
n = 1

for(n in 1:35){
  compliance_rate[r,2] <- length(which(EMAdata$id == id_names[n]))
  r <- r + 1
  n <- n + 1
  }
```

```{r}
#fill matrix col 3 with compliance rate
r = 1
n = 1

for (n in 1:35){
  compliance_rate[r, 3] <- compliance_rate[r, 2]/45
  r <- r + 1
  n <- n + 1
}
```

```{r}
#output min and max values
max <- which(compliance_rate == max(compliance_rate[,3]), arr.ind=TRUE)
min <- which(compliance_rate == min(compliance_rate[,3]), arr.ind=TRUE)

print(paste("Maximum compliance rate:", compliance_rate[max[1]]))
print(paste("Minimum compliance rate:", compliance_rate[min]))
```

### Time-Completed Distribution

```{r}
library(lubridate)

EMAdata %>%
  mutate(hour = hour(time_participant)) %>%
  ggplot(aes(x=hour)) + geom_bar() +
  ggtitle("Survey Timestamp Freq Dist")

#? - how to change x axis ticks+labels
```
## Score Summaries
```{r}
summarise_at(EMAdata, vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```
```{r}
#group by event
EMAdata%>%
  group_by(event) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```
```{r}
#group by enrollment
EMAdata%>%
  group_by(enrollmentgroup) %>%
  summarise_at(vars(positive_score, negative_score), list(mean=mean, sd=sd), na.rm=TRUE)
```
```{r}
#create grouped boxplot - positive

ggplot(EMAdata, aes(x=event, y=positive_score, fill=enrollmentgroup)) + 
  geom_boxplot() + xlab("Event") + ylab("Score") +
  ggtitle("PANAS Positive Score x Group")
```
```{r}
#create grouped boxplot - negative

ggplot(EMAdata, aes(x=event, y=negative_score, fill=enrollmentgroup)) + 
  geom_boxplot() + xlab("Event") + ylab("Score") +
  ggtitle("PANAS Negative Score x Group")
```

# Multilevel Model

## Preliminary Analysis
### Within-Days

```{r}
withinday <- EMAdata[,c('id', 'event', 'day', 'instance', 'enrollmentgroup', 'time_interval', 'positive_score', 'negative_score', "positive_SD","negative_SD")]
```

```{r}
#convert time_interval from h:m:s to minutes
withinday$time_interval_min <- hour(withinday$time_interval)*60 + minute(withinday$time_interval)

withinday <- withinday %>%
  relocate("time_interval_min", .after="time_interval")
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_SD)) + geom_point(size = 0.5) +
  ggtitle("Successive Difference - Negative")
```

#### Option 2
ASD = positive_SD[i] / time_interval[i]

```{r}
pos_ASD <- vector()

for(i in 1:1342){
  pos_ASD <- append(pos_ASD, c(withinday$positive_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(pos_ASD)
```

```{r}
#create positive ASD variable
withinday$positive_ASD <- pos_ASD
```

```{r}
neg_ASD <- vector()

for(i in 1:1342){
  neg_ASD <- append(neg_ASD, c(withinday$negative_SD[i] / withinday$time_interval_min[i]))
  
  i = i + 1
}

head(neg_ASD)
```

```{r}
#create negative ASD variable
withinday$negative_ASD <- neg_ASD
```

```{r}
# scatter plot of ASD x time interval (hour)
ggplot(withinday, aes(x = time_interval, y=positive_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Positive")
  
ggplot(withinday, aes(x = time_interval, y=negative_ASD)) + geom_point(size = 0.5) +
  ggtitle("Adj. Successive Difference - Negative")
```

### Between-Days
Average observations obtained within days to use as daily (mean) positive/negative affect score

```{r}
#initialize matrix
betweenday <- EMAdata[,c('id', 'event', 'day', 'instance', 'enrollmentgroup')]

betweenday$positive_mean <- 0
betweenday$negative_mean <- 0
```

```{r}
#reformat  
x = 1

for (x in 1:15){
  if(x != 1 & x != 6 & x != 11){
    betweenday <- betweenday[!(betweenday$instance == x),]
  }
  x = x + 1
}
```

```{r}
#fill positive mean

r = 1
mean_vec <- vector()

for(r in 1:35){ #number of ids
  
  d = 1
  
  for(d in 1:9){ #number of days
      
    x = 1
    score <- vector()
    
    score_i <- which(EMAdata$day == d & EMAdata$id == id_names[r])
      
    for(x in 1:length(score_i)){
      
      score <- append(score, c(EMAdata$positive_score[score_i[x]]))
      
    x = x + 1
    }
      
    score_mean <- round((mean(score, na.rm = TRUE)), digits = 2)
    if(is.nan(score_mean)) {score_mean <- NA} 
      
    mean_vec <- append(mean_vec, c(score_mean))
      
  d = d + 1
  }
r = r + 1
}

i = 1

for(i in 1:272){
  betweenday$positive_mean[i] <- mean_vec[i]
}
```

```{r}
#fill negative mean
i = 1
r = 1

mean_vec <- vector()

for(r in 1:35){
  
  d = 1
  
  for(d in 1:9){
      
    x = 1
    score <- vector()
    
    score_i <- which(EMAdata$day == d & EMAdata$id == id_names[r])
      
    for(x in 1:length(score_i)){
      
      score <- append(score, c(EMAdata$negative_score[score_i[x]]))
      
    x = x+1
    }
      
    score_mean <- round((mean(score, na.rm = TRUE)), digits = 2)
    if(is.nan(score_mean)) {score_mean <- NA} 
      
    mean_vec <- append(mean_vec, c(score_mean))
      
  d = d + 1
  }
r = r + 1
}

for(i in 1:272){
  betweenday$negative_mean[i] <- mean_vec[i]
}
```

```{r}
#error correction
betweenday$positive_mean[33] <- 27.4
betweenday$negative_mean[33] <- 22.0 

betweenday$positive_mean[66] <- 21.5
betweenday$negative_mean[66] <- 12.0

betweenday$positive_mean[104] <- 16.4
betweenday$negative_mean[104] <- 12.6

betweenday$positive_mean[156] <- 28.8
betweenday$negative_mean[156] <- 24.6

betweenday$positive_mean[160] <- 32.0
betweenday$negative_mean[160] <- 23.0

betweenday$positive_mean[161] <- 29.0
betweenday$negative_mean[161] <- 23.4

betweenday$positive_mean[162] <- 25.6
betweenday$negative_mean[162] <- 20.4

betweenday$positive_mean[271] <- 20.75
betweenday$negative_mean[271] <- 13.75
```

### Temporal Instability: MSSD

```{r}
# initialize LT-MSSD matrix

lt_mssd <- matrix(0, 35, 2)

colnames(lt_mssd) <- c("MSSD_pos", "MSSD_neg")
rownames(lt_mssd) <- id_names
```

```{r}
mssd(betweenday$positive_mean, group = betweenday$id, lag = 1, na.rm = TRUE)

for(i in 1:35){
  lt_mssd[1,1]
}
```

# Moore et al (2016) Analysis Plan

## Descriptives
### Group differences using t-tests (continuous) or chi-square (categorical)

## Mixed Model: Across events
```{r}
#create EMA_mm without "Group"

EMA_mm <- filter(EMAdata, enrollmentgroup != "Group")
betweenday_mm <- filter(betweenday, enrollmentgroup != "Group")
```

```{r}
#mixed models, repeated measures analysis of variance with REML
library(lmerTest)

lme_pos <- lmer(positive_mean ~ enrollmentgroup * event + (1|id), data=betweenday_mm)
  
  #fixed effects: enrollmentgroup, event, event*enrollmentgroup
  #random effects: ID

summary(lme_pos)
anova(lme_pos) 
```

```{r}
#mixed models, repeated measures analysis of variance with REML

lme_neg <- lmer(negative_mean ~ enrollmentgroup * event + (1|id), data=betweenday_mm)
  
  #fixed effects: enrollmentgroup, event, event*enrollmentgroup
  #random effects: ID

summary(lme_neg)
anova(lme_neg) 
```

### Create plots
```{r}
# Positive Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panaspos <- confint(lme_pos, method="boot", nsim=10)
ci_panaspos <- tibble::rownames_to_column(data.frame(ci_panaspos), "Term") 
colnames(ci_panaspos)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panaspos)

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday_mm, 
       aes(y=positive_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Timepoints") + 
  labs(y = "positive affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lme_pos)[1], #Regression Line (RL).
              slope=fixef(lme_pos)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panaspos$`CI 97.5%`[3], 
              slope=ci_panaspos$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panaspos$`CI 2.5%`[3], 
              slope=ci_panaspos$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL
  
panaspos_ggplot
```
```{r}
# Negative Affect plot
xlabels <- c("Baseline", "Session 4", "Post")

#create trendline
ci_panasneg <- confint(lme_neg, method="boot", nsim=10)
ci_panasneg <- tibble::rownames_to_column(data.frame(ci_panasneg), "Term") 
colnames(ci_panasneg)<- c("Term", "CI 2.5%", "CI 97.5%")
as.tibble(ci_panasneg)

#Plot model 
panasneg_ggplot<- ggplot(data= betweenday_mm, 
       aes(y=negative_mean, x=event, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Timepoints") + 
  labs(y = "negative affect score", x = "timepoint") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) +
  geom_abline(intercept = fixef(lme_neg)[1], #Regression Line (RL).
              slope=fixef(lme_neg)[2], col="black", linetype = "dashed") +                 
  geom_abline(intercept = ci_panasneg$`CI 97.5%`[3], 
              slope=ci_panasneg$`CI 97.5%`[4], col="grey", linetype = "dotted") + #Upper Bound of RL
  geom_abline(intercept = ci_panasneg$`CI 2.5%`[3], 
              slope=ci_panasneg$`CI 2.5%`[4], col="grey", linetype= "dotted") #Lower Bound of RL

panasneg_ggplot
```
## Mixed models - across days
```{r}
betweenday_mm$day <- as.factor(betweenday_mm$day)
levels(betweenday_mm$day) <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")
```

```{r}
#mixed models, repeated measures analysis of variance with REML

lme_pos <- lmer(positive_mean ~ enrollmentgroup * day + (1|id), data=betweenday_mm)
  
  #fixed effects: enrollmentgroup, event, event*enrollmentgroup
  #random effects: ID

anova(lme_pos) 

summary(lme_pos)
```


```{r}
#mixed models, repeated measures analysis of variance with REML

lme_neg <- lmer(negative_mean ~ enrollmentgroup * day + (1|id), data=betweenday_mm)
  
  #fixed effects: enrollmentgroup, event, event*enrollmentgroup
  #random effects: ID

anova(lme_neg) 

summary(lme_neg)
```

### Create plots
```{r}
# Positive Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday_mm, 
       aes(y=positive_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Days") + 
  labs(y = "positive affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels) 
 
panaspos_ggplot
```

```{r}
# Negative Affect plot
xlabels <- c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7", "Day 8", "Day 9")

#Plot model - positive
panaspos_ggplot<- ggplot(data= betweenday_mm, 
       aes(y=negative_mean, x=day, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Days") + 
  labs(y = "negative affect score", x = "day") +
  theme(legend.position = "none") + 
  scale_x_discrete(labels = xlabels)

panaspos_ggplot
```
# Mixed Models: across instances
```{r}
EMA_mm$instance <- as.factor(EMA_mm$instance)
levels(EMA_mm$instance) <- c("1", "2","3","4","5","6","7","8","9","10","11","12","13","14","15")
```

```{r}
#mixed models, repeated measures analysis of variance with REML
lme_pos <- lmer(positive_score ~ enrollmentgroup * instance + (1|id), data= EMA_mm)

anova(lme_pos) 
summary(lme_pos)
```
```{r}
#mixed models, repeated measures analysis of variance with REML
lme_neg <- lmer(negative_score ~ enrollmentgroup * instance + (1|id), data= EMA_mm)

anova(lme_neg) 
summary(lme_neg)
```

```{r}
# Positive Affect plot @ Baseline
EMA_baseline <- filter(EMA_mm, event == "Baseline")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_baseline, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Baseline") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```
```{r}
# Positive Affect plot @ Session 4
EMA_session4 <- filter(EMA_mm, event == "Session 4")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_session4, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Session4") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```
```{r}
# Positive Affect plot @ Post
EMA_post <- filter(EMA_mm, event == "Post")
  
#Plot model - positive
instance_pos<- ggplot(data= EMA_post, 
       aes(y=positive_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Positive Affect Across Instance @ Post") + 
  labs(y = "positive affect score", x = "instance") +
  theme(legend.position = "none") 

instance_pos
```


```{r}
# Negative Affect plot @ Baseline

instance_neg<- ggplot(data= EMA_baseline, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Baseline") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```
```{r}
# Negative Affect plot @ Session 4

instance_neg<- ggplot(data= EMA_session4, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Session 4") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```
```{r}
# Negative Affect plot @ Post

instance_neg<- ggplot(data= EMA_post, 
       aes(y=negative_score, x=instance, color=id, group=id)) + 
  geom_point() + geom_line() +
  ggtitle("Negative Affect Across Instance  @ Post") + 
  labs(y = "negative affect score", x = "instance") +
  theme(legend.position = "none") 

instance_neg
```


